---
title: "Worked examples"
output: 
# html_document:
#   keep_md: yes
# md_document:
#   variant: markdown_github
  github_document:
    dev: jpeg
    toc: true
    toc_depth: 2
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      fig.path='../output/figures/examples/examples_fig_',
                      #fig.height = 10,
                      cache.path = '../output/cache/examples/examples_cache_')
knitr::opts_knit$set(root.dir=normalizePath(".."))

set.seed(3128)

```

# Worked examples

```{r fxns, include=TRUE, message=FALSE}
require(pavo)
require(scatterplot3d)
require(gridExtra)
require(vegan)
require(MASS)
require(RColorBrewer)

# load aesthetic functions (plot, make colors transparent)
source('R/aesthetic.R')

# load function to convert JND to cartesian coordinates
source('R/jnd2xyz.R')

# load function for bootstrap
source('R/bootstrapcentroiddS.R')

# Distance matrix generator
distmat <- function(x){
  coldistres <- as.matrix(rbind(x[ ,c(1,2,3)], x[ ,c(2,1,3)]))
  uniquepatches <-  unique(c(coldistres[,1], coldistres[,2]))
  
  M <- matrix(nrow=length(uniquepatches), ncol=length(uniquepatches))
  
  rownames(M) <- colnames(M) <- uniquepatches
  
  M[coldistres[,1:2] ] <- coldistres[,3]
  M[coldistres[,2:1] ] <- coldistres[,3]
  
  class(M) <- 'numeric'
  M[is.na(M)] <- 0
  
  grouping <- as.factor(gsub('[0-9]','', rownames(M)))
  
  M <- as.dist(M)
  M
  }

#color palette
palette <- rcbalpha(0.8, 4, 'Set1')
```

## Example 1: Dichromatism. 

Reflectance data from four body regions of male and female _Ctenophorus ornatus_ (Whiting et al. 2015, Biol J Linn Soc). Labium, throat, tongue, and mouth-roof.

**Q:** Which body regions are sexually dichromatic?

Calculate deltaS according to conspecific (tetrachromatic) visual system

```{r message=FALSE, warning=FALSE}

specs <- list(all = as.rspec(read.csv('data/dichromatism/combined.csv'), interp = FALSE),
              lab = as.rspec(read.csv('data/dichromatism/lab.csv'), interp = FALSE),
              throat = as.rspec(read.csv('data/dichromatism/throat.csv'), interp = FALSE),
              roof = as.rspec(read.csv('data/dichromatism/roof.csv'), interp = FALSE),
              tongue = as.rspec(read.csv('data/dichromatism/tongue.csv'), interp = FALSE))

# Ctenophorus ornatus
liz_vis <- sensmodel(c(360, 440, 493, 571)) 
names(liz_vis) <- c('wl', 'u', 's', 'm', 'l')

models <- lapply(specs, vismodel, visual = liz_vis, relative = FALSE, qcatch='Qi')

spaces <- lapply(models, colspace)

deltaS <- lapply(models, coldist, achro = FALSE, n = c(1,1,3.5,6), 
                                  weber = 0.1, noise = "neural")
```

Visualise
```{r liz_spectcs, fig.width=6, fig.height=9, message=FALSE, warning=FALSE}
layout(matrix(c(1, 2, 3, 4, 5, 6, 7, 8), 4, 2, byrow = TRUE))

aggplot(specs[['lab']], by=gsub("[0-9].*","",names(specs[['lab']])), lwd=3, ylim=c(0,50))

scatterplot3d(spaces[['lab']][,c('x','y','z')],  
      bg=as.character(factor(gsub("[0-9].*","",names(specs[['lab']]))[-1], 
                             labels=palette[1:2])),
      box=FALSE, pch=21, cex.symbols=2, color=NA,
      x.ticklabs='', y.ticklabs='', z.ticklabs='', xlab='', ylab='', zlab='')


aggplot(specs[['throat']], by=gsub("[0-9].*","",names(specs[['throat']])), lwd=3, ylim=c(0,50))

scatterplot3d(spaces[['throat']][,c('x','y','z')],  
      bg=as.character(factor(gsub("[0-9].*","",names(specs[['throat']]))[-1],
                             labels=palette[1:2])),
      box=FALSE, pch=21, cex.symbols=2, color=NA,
      x.ticklabs='', y.ticklabs='', z.ticklabs='', xlab='', ylab='', zlab='')


aggplot(specs[['roof']], by=gsub("[0-9].*","",names(specs[['roof']])), lwd=3, ylim=c(0,50))

scatterplot3d(spaces[['roof']][,c('x','y','z')],  
      bg=as.character(factor(gsub("[0-9].*","",names(specs[['roof']]))[-1],
                             labels=palette[1:2])),
      box=FALSE, pch=21, cex.symbols=2, color=NA,
      x.ticklabs='', y.ticklabs='', z.ticklabs='', xlab='', ylab='', zlab='')


aggplot(specs[['tongue']], by=gsub("[0-9].*","",names(specs[['tongue']])), lwd=3, ylim=c(0,50))

scatterplot3d(spaces[['tongue']][,c('x','y','z')],  
      bg=as.character(factor(gsub("[0-9].*","",names(specs[['tongue']]))[-1],
                             labels=palette[1:2])),
      box=FALSE, pch=21, cex.symbols=2, color=NA,
      x.ticklabs='', y.ticklabs='', z.ticklabs='', xlab='', ylab='', zlab='')
```

**Step 1:** PERMANOVAs

```{r}

# Setup distance matrices & groupings for each body part
mat <- list(all = distmat(deltaS$all),
            lab = distmat(deltaS$lab),
            throat = distmat(deltaS$throat),
            roof = distmat(deltaS$roof),
            tongue = distmat(deltaS$tongue))

group <- list(all = paste0(substring(rownames(as.matrix(mat$all)), nchar(rownames(as.matrix(mat$all)))), substring(rownames(as.matrix(mat$all)), 1, 1)),
              lab = substring(rownames(as.matrix(mat$lab)), 1, 1),
              throat = substring(rownames(as.matrix(mat$throat)), 1, 1),
              roof = substring(rownames(as.matrix(mat$roof)), 1, 1),
              tongue = substring(rownames(as.matrix(mat$tongue)), 1, 1))

# Labium
adonis(mat$lab ~ group$lab)

# Mouth-roof
adonis(mat$roof ~ group$roof)

# Throat
adonis(mat$throat ~ group$throat)

# Tongue
adonis(mat$tongue ~ group$tongue)

```

**LESS SHITTY ALTERNATIVE?**

Combine all tests into one using sliding contrasts

```{r}

# Planned contrasts
cpatch <- factor(group$all)

levels(cpatch)  # First letter indicates body region, second indicates sex (see below for key)

contrasts(cpatch) <- contr.sdif(8)  # Sliding contrasts

# RM: maybe this works? 

contmat <- matrix(0, nrow=8, ncol=4)
contmat[1,1] <- contmat[3,2] <- contmat[5,3] <- contmat[7,4] <- 0.5
contmat[2,1] <- contmat[4,2] <- contmat[6,3] <- contmat[8,4] <- -0.5

contrasts(cpatch) <- contmat

# Design matrix without intercept
cgmmat <- model.matrix(~cpatch)[,-1]

# Run the model testing only contrasts of interest (m vs f body regions) i.e:
# HF vs HM (2-1), throat
# LM vs LF (4-3), labium
# RF vs RM (6-5), mouth-roof
# TF vs TM (8-7), tongue
#adonis2(mat$all ~ cgmmat[,'cpatch2-1'] + cgmmat[,'cpatch4-3'] + 
#          cgmmat[,'cpatch6-5'] + cgmmat[,'cpatch8-7'] , by = 'margin')

# cpatch1 = H
#       2 = L
#       3 = R
#       4 = T

adonis2(mat$all ~ cgmmat[,'cpatch1'] + cgmmat[,'cpatch2'] +
                  cgmmat[,'cpatch3'] + cgmmat[,'cpatch4'], by='margin')

```

TW: Yeah that's not right - see individual tests above. Should be labium (4-3) = distinct, throat (2-1) = distinct, mouth (6-5) = nope, tongue (8-7) = nope. Should 'sex' be split off as a factor?...

RM: my re-do results in only throat being significant... 

**Step 2:** Effect sizes.

Add grouping variable to raw models, then bootstrap centroids for statistically distinct patches, as identified in step 1 (labium & throat).

```{r}
# Groups
models$lab$group <- substring(rownames(models$lab), 1, 1)
models$throat$group <- substring(rownames(models$throat), 1, 1)
models$roof$group <- substring(rownames(models$roof), 1, 1)
models$tongue$group <- substring(rownames(models$tongue), 1, 1)

# labium
laboot <- bootcentroidDS(models$lab[,1:4], models$lab$group, 
                         n=c(1,1,3.5,6), weber=0.1, qcatch='Qi', achro=FALSE)

# throat
thboot <- bootcentroidDS(models$throat[,1:4], models$throat$group, 
                         n=c(1,1,3.5,6), weber=0.1, qcatch='Qi', achro=FALSE)

# roof
roboot <- bootcentroidDS(models$roof[,1:4], models$roof$group, 
                         n=c(1,1,3.5,6), weber=0.1, qcatch='Qi', achro=FALSE)

# tongue
toboot <- bootcentroidDS(models$tongue[,1:4], models$tongue$group, 
                         n=c(1,1,3.5,6), weber=0.1, qcatch='Qi', achro=FALSE)

```

```{r}
bootres <- rbind(laboot, thboot, roboot, toboot)
rownames(bootres) <- c('Labium', 'Throat', 'Roof', 'Tongue')

bootres

plot(bootres[,1], xlim=c(0.5, 4.5), ylim=c(0, 2.2), pch=21, bg=1, cex=2, xaxt='n', xlab='Centroid comparison', ylab='Chromatic contrast (JND)')

abline(h=1, lty=3, lwd=2)
segments(1:4, bootres[,2], 1:4, bootres[,3], lwd=2)

axis(1, at=1:4, labels=rownames(bootres))

```

So lab's & throats are statistically distinct, but fall below threshold.

plots for publication:

```{r}
pdf(height=10*1.3, width=8*1.3, pointsize = 12*1.3*1.2, file='figures/lizards.pdf')

oma <- par()$mar

oma[3] <- 1.1
oma[1] <- 4.1

layout(cbind(matrix(c(1, 2, 3, 4, 5, 6, 7, 8), 4, 2, byrow = TRUE), rep(9,4)))

par(mar=oma)

aggplot(specs[['lab']], by=gsub("[0-9].*","",names(specs[['lab']])), lwd=3, ylim=c(0,50))
text(x=grconvertX(0.05,"npc"), y=grconvertY(0.95, "npc"), cex=1.5, "A") 

scatterplot3d(spaces[['lab']][,c('x','y','z')],  
      bg=as.character(factor(gsub("[0-9].*","",names(specs[['lab']]))[-1], 
                             labels=palette[1:2])),
      box=FALSE, pch=21, cex.symbols=2, color=NA,
      x.ticklabs='', y.ticklabs='', z.ticklabs='', xlab='', ylab='', zlab='',
      mar=c(4.5,0,0,0))

par(mar=oma)

aggplot(specs[['throat']], by=gsub("[0-9].*","",names(specs[['throat']])), lwd=3, ylim=c(0,50))
text(x=grconvertX(0.05,"npc"), y=grconvertY(0.95, "npc"), cex=1.5, "B") 

scatterplot3d(spaces[['throat']][,c('x','y','z')],  
      bg=as.character(factor(gsub("[0-9].*","",names(specs[['throat']]))[-1],
                             labels=palette[1:2])),
      box=FALSE, pch=21, cex.symbols=2, color=NA,
      x.ticklabs='', y.ticklabs='', z.ticklabs='', xlab='', ylab='', zlab='',
      mar=c(4.5,0,0,0))

par(mar=oma)

aggplot(specs[['roof']], by=gsub("[0-9].*","",names(specs[['roof']])), lwd=3, ylim=c(0,50))
text(x=grconvertX(0.05,"npc"), y=grconvertY(0.95, "npc"), cex=1.5, "C") 

scatterplot3d(spaces[['roof']][,c('x','y','z')],  
      bg=as.character(factor(gsub("[0-9].*","",names(specs[['roof']]))[-1],
                             labels=palette[1:2])),
      box=FALSE, pch=21, cex.symbols=2, color=NA,
      x.ticklabs='', y.ticklabs='', z.ticklabs='', xlab='', ylab='', zlab='',
      mar=c(4.5,0,0,0))

par(mar=oma)

aggplot(specs[['tongue']], by=gsub("[0-9].*","",names(specs[['tongue']])), lwd=3, ylim=c(0,50))
text(x=grconvertX(0.05,"npc"), y=grconvertY(0.95, "npc"), cex=1.5, "D") 

scatterplot3d(spaces[['tongue']][,c('x','y','z')],  
      bg=as.character(factor(gsub("[0-9].*","",names(specs[['tongue']]))[-1],
                             labels=palette[1:2])),
      box=FALSE, pch=21, cex.symbols=2, color=NA,
      x.ticklabs='', y.ticklabs='', z.ticklabs='', xlab='', ylab='', zlab='',
      mar=c(4.5,0,0,0))

par(mar=oma)
par(mar=c(4,4,1,1)+0.1)

plot(rev(bootres[,1]), 1:4, ylim=c(0.8, 4.2), xlim=c(0, 2.5), pch=21, bg=1, cex=2, yaxt='n', ylab='Body patch', xlab='Chromatic contrast (JND)')
axis(2, at=1:4, labels=rev(rownames(bootres)), las=1)

abline(v=1, lty=3, lwd=2)
segments(rev(bootres[,2]), 1:4, rev(bootres[,3]), 1:4, lwd=2)
text(x=grconvertX(0.05,"npc"), y=grconvertY(0.99, "npc"), cex=1.5, "E") 

dev.off()

```

```{r}

rm(list = setdiff(ls(), lsf.str()))

```

## Example 2: Mimicry. 

Reflectance data from colour-polymorphic female spiders _Gasteracantha fornicata_, and sympatic flowers from Qld, Australia. (W = white morph, Y = yellow morph, F = flowers)

So three groups, with **two Q's:** 

**(1)** Are spiders actually polymorphic (to prey), as naively seems to be the case (for humans)?

**(2)** Do spiders (of each morph) resemble sympatric flowers?

Calculate deltaS (JNDs) according to a honeybee

```{r}
specs <- as.rspec(read.csv('data/mimicry/flowers_spiders.csv'), interp = FALSE)

models <- vismodel(specs, visual = 'apis', relative = FALSE)

spaces <- colspace(models)

#deltaS <- coldist(models, achro = FALSE, n1 = 1, n2 = 0.471, n3 = 4.412, v = 0.13)
deltaS <- coldist(models, achro=FALSE, n=c(1,0.5,4.4), weber=0.13)

models$group <- substring(rownames(models), 1, 1)
bootcentroidDS(models[, 1:3], models$group, n=c(1,0.5,4.4), weber=0.13, achro=FALSE, qcatch='Qi')

# Contrast labels
deltaS$comparison[grepl('W_', deltaS$patch1) & grepl('W_', deltaS$patch2)] <- 'intra.W'
deltaS$comparison[grepl('Y_', deltaS$patch1) & grepl('Y_', deltaS$patch2)] <- 'intra.Y'
deltaS$comparison[grepl('F_', deltaS$patch1) & grepl('F_', deltaS$patch2)] <- 'intra.F'
deltaS$comparison[grepl('Y_', deltaS$patch1) & grepl('W_', deltaS$patch2)] <- 'inter.WY'
deltaS$comparison[grepl('W_', deltaS$patch1) & grepl('F_', deltaS$patch2)] <- 'inter.WF'
deltaS$comparison[grepl('Y_', deltaS$patch1) & grepl('F_', deltaS$patch2)] <- 'inter.YF'

```

Visualise.

```{r mimic_triplot, fig.width=6, fig.height=6}
# specs
aggplot(specs, by=substr(names(specs), 1,1)[-1], shadecol=c('darkgoldenrod1', 'darkgrey', 'forestgreen'), alpha=c(0.8,0.8,0.2), lcol=1, lwd=2)

aggplot(specs[,grep('F', names(specs), invert=T)], by=substr(grep('F', names(specs), invert=TRUE, value=TRUE), 1,1)[-1], shadecol=c('goldenrod1', 'darkgrey'), alpha=0.5, lcol=1, lwd=2, ylim=c(0,100))

# Max triangle
plot(spaces, pch=21, cex=2, bg=as.character(factor(substr(rownames(spaces), 1, 1), labels=c('forestgreen', 'darkgrey', 'darkgoldenrod1'))))

```

**PERMANOVA**
```{r}

# Set up distance matrices & groupings for focal comparisons 
mat <- list(all = distmat(deltaS),
            WY = distmat(subset(deltaS, !(comparison %in% c('intra.F', 'inter.WF', 'inter.YF')))),
            WF = distmat(subset(deltaS, !(comparison %in% c('intra.Y', 'inter.WY', 'inter.YF')))),
            YF = distmat(subset(deltaS, !(comparison %in% c('intra.W', 'inter.WY', 'inter.WF'))))
            )
group <- list(all = substring(rownames(as.matrix(mat$all)), 1, 1),
              WY = substring(rownames(as.matrix(mat$WY)), 1, 1),
              WF = substring(rownames(as.matrix(mat$WF)), 1, 1),
              YF = substring(rownames(as.matrix(mat$YF)), 1, 1))

# test if W and Y are different
adonis(mat$WY ~ group$WY)

# test if W and Y are different from F
# Use customized contrasts to test a priori hypotheses together
cgroups <- factor(group$all)

contrasts(cgroups) <- contr.treatment(3)

contrasts(cgroups) <- cbind(
  c(-1,1,0),
  c(-1,0,1)
)

# Create design matrix, without intercept
cgmmat <- model.matrix(~cgroups)[,-1]
colnames(cgmmat) <- c('W','Y')

# Run the model testing only specified contrasts
adonis2(mat$all ~ cgmmat[,'W'] + cgmmat[,'Y'], by='margin')
```

Note that the degrees of freedom, sums of squares, and R2 are the same between the two "full" models, but the specified contrasts let us test both hypotheses together. They would also be the same if you did adonis(mat$all ~ cgroups) (but you'd only get one P value and R2 representing the combined effects of both sets of contrasts)

**Effect sizes**
```{r}

models$group <- substring(rownames(models), 1, 1)
spiderboot <- bootcentroidDS(models[, 1:3], models$group, n=c(1, 0.5, 4.4), weber = 0.13, qcatch='Qi', achro=FALSE)

spiderboot

plot(spiderboot[,1], xlim=c(0.5, 3.5), ylim=c(0, 3), pch=21, bg=1, cex=2, xaxt='n', ylab='Chromatic distance (JND)', xlab='Centroid comparison')
axis(1, at=1:3, labels = c('Flower-White', 'Flower-Yellow', 'White-Yellow'))
segments(1:3, spiderboot[,2], 1:3, spiderboot[,3], lwd=2)
abline(h=1, lty=3, lwd=2)
abline(h=0.3, lty=3, lwd=2, col='darkgrey')
```

So the RN threshold for honeybees can be pretty damn low (0.3 JNDs, Dyer & Neumeyer 2005), but is variable depending on testing conditions, past experience etc. These would suggest that everying's (on average) perceptably distinct, but probably tough (depending on experience etc.). 

plot for paper:

```{r}
# transparent colors
spidercol <- rgb(data.frame(t(col2rgb(c('forestgreen', 'darkgrey', 'darkgoldenrod1'),
                                      alpha=F))/255), alpha=0.8)

pdf(height=8*1.3, width=4*1.3, pointsize = 12, file='figures/spiders1.pdf')

par(mfrow=c(3,1), mar=c(5.1, 4.1, 2.1, 2.1))

aggplot(specs[,grep('F', names(specs), invert=T)], by=substr(grep('F', names(specs), invert=TRUE, value=TRUE), 1,1)[-1], shadecol=c('goldenrod1', 'darkgrey'), alpha=0.5, lcol=1, lwd=2, ylim=c(0,100))
text(x=grconvertX(0.05,"npc"), y=grconvertY(0.95, "npc"), cex=1.5, "A") 

plot(spaces, pch=21, cex=2, 
     margins=c(0,0,3,0),
     bg=as.character(factor(substr(rownames(spaces), 1, 1), labels=spidercol)),
     col=NA, achrocol='black')
text(x=grconvertX(0.05,"npc"), y=grconvertY(0.95, "npc"), cex=1.5, "B") 


plot(spiderboot[,1], xlim=c(0.5, 3.5), ylim=c(0, 3), pch=21, bg=1, cex=2, xaxt='n', ylab='Chromatic distance (JND)', xlab='Centroid comparison')
axis(1, at=1:3, labels = c('Flower-White', 'Flower-Yellow', 'White-Yellow'))
abline(h=1, lty=3, lwd=2)
abline(h=0.3, lty=3, lwd=2, col='darkgrey')
segments(1:3, spiderboot[,2], 1:3, spiderboot[,3], lwd=2)
text(x=grconvertX(0.05,"npc"), y=grconvertY(0.95, "npc"), cex=1.5, "C") 

dev.off()



pdf(height=8*1.3, width=4*1.3, pointsize = 12, file='figures/spiders2.pdf')

par(mfrow=c(3,1), mar=c(5.1, 4.1, 2.1, 2.1))

aggplot(specs, by=substr(names(specs), 1,1)[-1], shadecol=c('darkgoldenrod1', 'darkgrey', 'forestgreen'), alpha=c(0.8,0.8,0.2), lcol=1, lwd=2, ylim=c(0,100))

plot(spaces, pch=21, cex=2, 
     margins=c(0,0,3,0),
     bg=as.character(factor(substr(rownames(spaces), 1, 1), labels=spidercol)),
     col=NA, achrocol='black')


plot(spiderboot[,1], xlim=c(0.5, 3.5), ylim=c(0, 3), pch=21, bg=1, cex=2, xaxt='n', ylab='Chromatic contrast (JND)', xlab='Centroid comparison')
axis(1, at=1:3, labels = c('Flower-White', 'Flower-Yellow', 'White-Yellow'))
abline(h=1, lty=3, lwd=2)
abline(h=0.3, lty=3, lwd=2, col='darkgrey')
segments(1:3, spiderboot[,2], 1:3, spiderboot[,3], lwd=2)

dev.off()

```

```{r}
sessionInfo()
```
