---
title: "Simulations - threshold"
output: 
# html_document:
#   keep_md: yes
# md_document:
#   variant: markdown_github
  github_document:
    dev: jpeg
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      fig.path='../output/figures/final/final_threshold_fig', 
                      #fig.height = 10,
                      cache.path = '../output/cache/final/final_threshold_cache_')
knitr::opts_knit$set(root.dir=normalizePath(".."))

set.seed(6210)
```


```{r fxns, include=TRUE}
require(pavo)
require(vegan)
require(RColorBrewer)

# load aesthetic functions (plot, make colors transparent)
source('R/aesthetic.R')

# load function to convert JND to cartesian coordinates
source('R/jnd2xyz.R')

# load simulation and analysis functions
source('R/simfoos.R')
source('R/simanalysis.R')
source('R/pausemcl.R')
```

# Threshold scenario: high within-group variability, centroid distance ~1JND

Generate data
```{r coldistcache3, cache=TRUE}
simN <- 50

simulatedata3a <- replicate(500, 
# this gave some interesting results we might want to revisit.
#                  simdich(N=simN, sgsqsrate=1, sdmeanratio=FALSE, multiplier=c(0.95, 1.05)), 
                  simdich(N=simN, sgsqsrate=5, sdmeanratio=FALSE, multiplier=c(0.97, 1.03)),
                  simplify=FALSE)

simulatedata3b <- replicate(500, 
                  simdich(N=simN, sgsqsrate=5, sdmeanratio=FALSE, multiplier=c(0.97, 1.03)),
                  simplify=FALSE)

simulatedata3 <- c(simulatedata3a, simulatedata3b)
rm(simulatedata3a, simulatedata3b)
# we need to add the reference points since we didn't run vismodel()
#apply(do.call(rbind, lapply(simulatedata, function(x) apply(x, 2, max))), 2, max)
#data(sicalis)
#rfs <- attr(vismodel(sicalis, visual='star',relative=FALSE), 'resrefs')

rfs <- 
matrix(c(100,01,01,01,
         01,100,01,01,
         01,01,100,01,
         01,01,01,100,
         50,50,50,50
         ), ncol=4, byrow=TRUE)
rownames(rfs) <- c('refforjnd2xyz.u','refforjnd2xyz.s','refforjnd2xyz.m','refforjnd2xyz.l', 'refforjnd2xyz.acent')
colnames(rfs) <- c('u','s','m','l')

simulatedata3 <- lapply(simulatedata3, 'attr<-', which='resrefs', value=rfs)


simulatecoldist3a <- pausemcl(simulatedata3[1:500], function(x) {
  Y <- suppressWarnings(coldist(x, achro=FALSE, qcatch='Qi'))
  Y$comparison <- NA
  Y$comparison[grepl('A', Y$patch1) & grepl('A', Y$patch2)] <- 'intra.A'
  Y$comparison[grepl('B', Y$patch1) & grepl('B', Y$patch2)] <- 'intra.B'
  Y$comparison[grepl('A', Y$patch1) & grepl('B', Y$patch2)] <- 'inter'
  Y
  } )

simulatecoldist3b <- pausemcl(simulatedata3[501:1000], function(x) {
  Y <- suppressWarnings(coldist(x, achro=FALSE, qcatch='Qi'))
  Y$comparison <- NA
  Y$comparison[grepl('A', Y$patch1) & grepl('A', Y$patch2)] <- 'intra.A'
  Y$comparison[grepl('B', Y$patch1) & grepl('B', Y$patch2)] <- 'intra.B'
  Y$comparison[grepl('A', Y$patch1) & grepl('B', Y$patch2)] <- 'inter'
  Y
  } )

simulatecoldist3 <- c(simulatecoldist3a, simulatecoldist3b)
rm(simulatecoldist3a, simulatecoldist3b)
```

Validating simulations:

```{r, echo=FALSE, dependson='coldistcache3'}
par(mfrow=c(2,2))
# GROUP A MEANS
empmeans <- do.call(rbind, lapply(simulatedata3, function(x) colMeans(x[1:simN,])))
simmeans <- do.call(rbind, lapply(simulatedata3, function(x) attr(x,'simpar')[1,]))

plot(empmeans[,'u']~simmeans[,'u'], pch=19, 
     col=rgb(data.frame(t(col2rgb('violet', alpha=F)))/255, alpha=0.2), 
     ylab='empirical',xlab='simulated', 
     xlim=range(rbind(empmeans,simmeans)), 
     ylim=range(rbind(empmeans,simmeans)),
     main='mean group A')
points(empmeans[,'s']~simmeans[,'s'], pch=19, col=rgb(data.frame(t(col2rgb('blue', alpha=F)))/255, alpha=0.2))
points(empmeans[,'m']~simmeans[,'m'], pch=19, col=rgb(data.frame(t(col2rgb('darkgreen', alpha=F)))/255, alpha=0.2))
points(empmeans[,'l']~simmeans[,'l'], pch=19, col=rgb(data.frame(t(col2rgb('tomato', alpha=F)))/255, alpha=0.2))

abline(0,1)

# GROUP B MEANS
empmeans <- do.call(rbind, lapply(simulatedata3, function(x) colMeans(x[(simN+1):(simN*2),])))
simmeans <- do.call(rbind, lapply(simulatedata3, function(x) attr(x,'simpar')[2,]))

plot(empmeans[,'u']~simmeans[,'u'], pch=19, 
     col=rgb(data.frame(t(col2rgb('violet', alpha=F)))/255, alpha=0.2), 
     ylab='empirical',xlab='simulated', 
     xlim=range(rbind(empmeans,simmeans)), 
     ylim=range(rbind(empmeans,simmeans)),
     main = 'mean group B')
points(empmeans[,'s']~simmeans[,'s'], pch=19, col=rgb(data.frame(t(col2rgb('blue', alpha=F)))/255, alpha=0.2))
points(empmeans[,'m']~simmeans[,'m'], pch=19, col=rgb(data.frame(t(col2rgb('darkgreen', alpha=F)))/255, alpha=0.2))
points(empmeans[,'l']~simmeans[,'l'], pch=19, col=rgb(data.frame(t(col2rgb('tomato', alpha=F)))/255, alpha=0.2))

abline(0,1)
# GROUP A SD
empmeans <- do.call(rbind, lapply(simulatedata3, function(x) apply(x[1:simN,], 2, sd)))
simmeans <- do.call(rbind, lapply(simulatedata3, function(x) attr(x,'simpar')[3,]))


plot(empmeans[,'u']~simmeans[,'u'], pch=19, 
     col=rgb(data.frame(t(col2rgb('violet', alpha=F)))/255, alpha=0.2), 
     ylab='empirical',xlab='simulated', 
     xlim=range(rbind(empmeans,simmeans)), 
     ylim=range(rbind(empmeans,simmeans)),
     main='SD group A')
points(empmeans[,'s']~simmeans[,'s'], pch=19, col=rgb(data.frame(t(col2rgb('blue', alpha=F)))/255, alpha=0.2))
points(empmeans[,'m']~simmeans[,'m'], pch=19, col=rgb(data.frame(t(col2rgb('darkgreen', alpha=F)))/255, alpha=0.2))
points(empmeans[,'l']~simmeans[,'l'], pch=19, col=rgb(data.frame(t(col2rgb('tomato', alpha=F)))/255, alpha=0.2))

abline(0,1)
# GROUP B SD
empmeans <- do.call(rbind, lapply(simulatedata3, function(x) apply(x[(simN+1):(simN*2),], 2, sd)))
simmeans <- do.call(rbind, lapply(simulatedata3, function(x) attr(x,'simpar')[3,]))

plot(empmeans[,'u']~simmeans[,'u'], pch=19, 
     col=rgb(data.frame(t(col2rgb('violet', alpha=F)))/255, alpha=0.2), 
     ylab='empirical',xlab='simulated', 
     xlim=range(rbind(empmeans,simmeans)), 
     ylim=range(rbind(empmeans,simmeans)),
     main='SD group B')
points(empmeans[,'s']~simmeans[,'s'], pch=19, col=rgb(data.frame(t(col2rgb('blue', alpha=F)))/255, alpha=0.2))
points(empmeans[,'m']~simmeans[,'m'], pch=19, col=rgb(data.frame(t(col2rgb('darkgreen', alpha=F)))/255, alpha=0.2))
points(empmeans[,'l']~simmeans[,'l'], pch=19, col=rgb(data.frame(t(col2rgb('tomato', alpha=F)))/255, alpha=0.2))

abline(0,1)
``` 

Verifying that values obtained in the simulation (empirical) are close to what we wanted to simulate (simulated) for the four cones (violet, blue, green, red)

```{r histograms, dependson='coldistcache3', echo=FALSE}
gmeansT <- lapply(simulatecoldist3, function(x) tapply(x$dS, x$comparison, mean))
gmeansT <- do.call(rbind, gmeansT)

# Get centroids and within/among mean distances; how manny above 1JND
centdistT <- unlist(pausemcl(simulatedata3, centroidist))

intradistT <- rowMeans(gmeansT[, -1])
interdistT <- gmeansT[,"inter"]

centroidPT <- centdistT > 1

# Plot distributions
palette <- rcbalpha(0.6, 3, 'Set1')

h1 <- hist(interdistT, breaks=seq(0, ceiling(max(interdistT))+1,by=0.5), plot=F)
h2 <- hist(intradistT, breaks=seq(0, ceiling(max(intradistT))+1,by=0.5), plot=F)
h3 <- hist(centdistT, breaks=seq(0, ceiling(max(centdistT))+1,by=0.5), plot=F)

par(mar=c(4,4.5,1.5,1.5))

plot(c(0,h1$breaks), c(0,h1$counts,0), type='s',col=palette[1], lwd=5,
     ylab='Frequency', xlab='mean distance (JND)',
     ylim=range(c(h1$counts,h2$counts,h3$counts))* c(0,1.1))
lines(c(0,h2$breaks), c(0,h2$counts,0), col=palette[2], type='s', lwd=5)
lines(c(0,h3$breaks), c(0,h3$counts,0), col=palette[3], type='s', lwd=5)

legend('topright', col=palette, c('between-group  ','within-group', 'centroid'), 
       pch=15, pt.cex=2.5, bty='n')

```

mean centroid distance of `r mean(centdistT)`, quantiles of `r quantile(centdistT, c(0.025, 0.975))`

## Running analyses

```{r adoniscache3, results='hide', cache=TRUE, dependson='coldistcache3'}
adonissim3a <- pausemcl(simulatecoldist3[1:500], adoniscoldist )
Sys.sleep(10)
adonissim3b <- pausemcl(simulatecoldist3[501:1000], adoniscoldist )
adonissim3 <- c(adonissim3a, adonissim3b)
vovsim3 <- pausemcl(simulatedata3, voloverlaptest )

scd23 <- lapply(simulatecoldist3,'[', ,1:3, drop=FALSE)
for(i in 1:length(scd23)){
  attr(scd23[[i]], 'resrefs') <- attr(simulatecoldist3[[i]],'resrefs')
  attr(scd23[[i]], 'conenumb') <- attr(simulatecoldist3[[i]],'conenumb')
  }
```

```{r pyke3cache, cache=TRUE, dependson='coldistcache3'}
pykesim3 <- lapply(scd23, jnd2xyz)
pykelm3 <- lapply(pykesim3, function(x) lm(as.matrix(x) ~ rep(c('gA','gB'), each=50)))
pykemanova3 <- lapply(pykelm3, function(x) summary(manova(x)))
vovpyke3 <- pausemcl(pykesim3, function(x)
  voloverlap(x[1:simN,], x[(simN+1):(simN*2), ]) )

gc(verbose=FALSE)
```


```{r, echo=FALSE, fig.width=7, fig.height=4.5, dependson=c('coldistcache3', 'adoniscache3', 'pyke3cache')}
# Pyke MANOVA: P-values, which significant
manovaPvalT <- unlist(lapply(pykemanova3, function(x) x$stats[1,'Pr(>F)']))
manovaPT <- manovaPvalT < 0.05

# Adonis: P-values, which significant, R-squared
adonisPvalT <- unlist(lapply(adonissim3, function(x) x$aov.tab$'Pr(>F)'[1]))
adonisPT <- adonisPvalT < 0.05
adonisR2T <- unlist(lapply(adonissim3, function(x) x$aov.tab$'R2'[1])) * 100

# disagreement in significance between Pyke and Adonis
disagreementT <- !manovaPT == adonisPT

# Volume overlap
overlapT <- unlist(lapply(vovsim3, '[','vboth')) * 100
overlapykeT <- unlist(lapply(vovpyke3, '[','vboth')) * 100

# Mahalanobis distance between centroids
mahdT <- unlist(lapply(simulatedata3, function(x) sqrt(mahalanobis(colMeans(x[1:50,]), colMeans(x[51:100,]), cov(x[1:50,]) ))))


sigpalT <- as.character(factor(adonisPT, labels=palette[1:2]))
sigpalT <- as.character(factor(paste(adonisPT, centroidPT), 
          levels=c("FALSE FALSE", "FALSE TRUE", "TRUE FALSE", "TRUE TRUE"), 
          labels=rcbalpha(0.8, 6, 'RdBu')[c(3,1,6,4)] ))
#sigpalT[sigpalT=="#FDDBC7CC"] <- "#D1E5F0CC"
```

##Visualizing results
  
color legend:

* dark colors: methods disagree (BAD)
* light colors: methods agree (GOOD)
  
* light blue: adonis and centroid distance > 1 (GOOD)
* dark blue: adonis significant, centroid distance < 1 (BAD)
* dark red: adonis non-significant, centroid distance > 1 (BAD)
* light red: adonis and centroid distance < 1 (GOOD)
```{r, echo=FALSE, fig.width=7, fig.height=7, dependson=c('coldistcache3', 'adoniscache3', 'pyke3cache')}

par(mfrow=c(3,3))

plot(interdistT~intradistT, ylab='mean between-group distance (JND)', xlab='mean within-group distance (JND)', pch=19, col=sigpalT, ylim=range(c(interdistT,intradistT))*1.1, xlim=range(c(interdistT,intradistT))*1.1)
abline(0,1, lty=3)

# Intradist plot is essentially identical (see histogram)
#plot(intradist~centdist, ylab='mean between-group distance (JND)', xlab='centroid distance (JND)', pch=19, col=rgb(0,0,0,0.4), ylim=c(0,5.5), xlim=c(0,5.5))
#abline(0,1, lty=3)

plot(adonisR2T~interdistT, ylab='Rsquared from PERMANOVA (%)', xlab='mean between-group distance (JND)', log='xy', pch=19, col=sigpalT)

plot(adonisR2T~intradistT, ylab='Rsquared from PERMANOVA (%)', xlab='mean within-group  distance (JND)', log='xy', pch=19, col=sigpalT)

plot(centdistT~intradistT, xlab='mean within-group distance (JND)', ylab='centroid distance (JND)', pch=19, col=sigpalT, ylim=range(c(intradistT,centdistT))*1.1, xlim=range(c(intradistT,centdistT))*1.1, log='xy')
abline(0,1, lty=3)
abline(h=1, lty=2)
abline(v=1, lty=2)
abline(-0.5, 1, lty=3, col='grey', lwd=2)


plot(adonisR2T~centdistT, ylab='Rsquared from PERMANOVA (%)', xlab='centroid distance (JND)', log='xy', pch=19, col=sigpalT)

plot(adonisR2T~overlapT, ylab='Rsquared from PERMANOVA (%)', xlab='color volume overlap',  pch=19, col=sigpalT, log='y')

plot(interdistT,overlapT,  pch=19, col=sigpalT, 
     xlab='mean between-group distance (JND)', ylab='color volume overlap')

plot(intradistT, overlapT,  pch=19, col=sigpalT, 
     xlab='mean within-group distance (JND)', ylab='color volume overlap')

plot(centdistT, overlapT,  pch=19, col=sigpalT, 
     xlab='centroid distance (JND)', ylab='color volume overlap')


#yesnohist(interdist, xlab='mean between-group distance (JND)')
#yesnohist(intradist, xlab='mean within-group distance (JND)')
#yesnohist(centdist, xlab='centroid distance (JND)')


# boxplot(intradistM~I(adonisPT < 0.05), 
#         xlab='PERMANOVA test significant?', ylab='mean within-group distance (JND)', 
#         boxwex=0.3, col=grey(0.7), pch=19, log='y')
```

```{r}
sessionInfo()
```

plots for publication:

```{r plotsforpub}
plotrange <- function(x, log=TRUE){
  res <- range(x)
  res[1] <- floor(res[1])
  res[2] <- ceiling(res[2])
  if(log && res[1] == 0)
    res[1] <- range(x)[1]*0.8
  
  if(log && res[2] < 10)
    res[2] <- 10
  
  res
}

######################
# RESULTS FROM SIM 2 #
######################
pdf(height=4*1.3, width=7*1.3, file='figures/threshold_1.pdf')
par(mfrow=c(2,3), mar=c(4,5,1,1))

plot(centdistT~intradistT, 
     xlab='mean within-group distance (JND)\n ', ylab=' \ncentroid distance (JND)', 
     ylim=plotrange(c(intradistT,centdistT)), xlim=plotrange(c(intradistT,centdistT)), 
     pch=21, bg=sigpalT, col=NA, log='xy', yaxt='n', xaxt='n')
axis(1, at=c(0.1, 1, 10), labels=c(0.1, 1, 10))
axis(1, at=c(seq(0.2,0.9, by=0.1), seq(2,9, by=1)), tcl=par("tcl")*0.5, labels=FALSE)
axis(2, at=c(0.1, 1, 10), labels=c(0.1, 1, 10))
axis(2, at=c(seq(0.2,0.9, by=0.1), seq(2,9, by=1)), tcl=par("tcl")*0.5, labels=FALSE)

abline(0,1, lty=2)
abline(h=1, lty=3)
abline(v=1, lty=3)
abline(-0.5, 1, lty=5, col='grey')

#the grey line represents an intercept of 1/sqrt(4) so MahD = 1

legend('topleft', pch=21, cex=0.9, bty='n', col=NA, pt.bg=rcbalpha(1, 6, 'RdBu')[c(1,3,6,4)],
       legend=c('p > 0.05, JND > 1',
                'p > 0.05, JND < 1',
                'p < 0.05, JND < 1',
                'p < 0.05, JND > 1'))

plot(adonisR2T~centdistT, 
     ylab=expression(paste(R^2,' (%)')), xlab='centroid distance (JND)\n ', 
     ylim=plotrange(adonisR2T), xlim=plotrange(centdistT),
     pch=21, bg=sigpalT, col=NA, log='xy', yaxt='n', xaxt='n')
#axis(2, at=c(0.05, 0.5, 5, 50), labels=c(0.05, 0.5, 5, 50))
axis(2, at=c(0.01, 0.1, 1, 10,100), labels=c(0.01, 0.1, 1, 10,100))
#axis(2, at=c(0.06,0.07,0.08,0.09, 0.2, 0.3, 0.4, 0.6, 0.7, 0.8, 0.9, 2, 3, 4, 6, 7, 8, 9, 20, 30, 40), tcl=par("tcl")*0.5, labels=FALSE)
axis(2, at=c(seq(0.02,0.09,by=0.01), seq(0.2,0.9,by=0.1), seq(2,9,by=1), seq(20,90,by=10)), tcl=par("tcl")*0.5, labels=FALSE)
#axis(2, at=c(0.1, 1, 10), tcl=par("tcl")*1, labels=FALSE)
axis(1, at=c(0.1, 1, 10), labels=c(0.1, 1, 10))
axis(1, at=c(seq(0.2,0.9, by=0.1), seq(2,9, by=1)), tcl=par("tcl")*0.5, labels=FALSE)

abline(v=1, lty=3)
abline(h=3, lty=3)

plot(adonisR2T~overlapT, 
     ylab=expression(paste(R^2,' (%)')), xlab='color volume overlap (%)\n ',  
     ylim=c(0.01, 100), xlim=plotrange(overlapT, log=FALSE),
     pch=21, bg=sigpalT, col=NA, log='y', yaxt='n', xaxt='n')
axis(1, at=c(0,20,40,60))
#axis(2, at=c(0.05, 0.5, 5, 50), labels=c(0.05, 0.5, 5, 50))
axis(2, at=c(0.01, 0.1, 1, 10,100), labels=c(0.01, 0.1, 1, 10,100))
#axis(2, at=c(0.06,0.07,0.08,0.09, 0.2, 0.3, 0.4, 0.6, 0.7, 0.8, 0.9, 2, 3, 4, 6, 7, 8, 9, 20, 30, 40), tcl=par("tcl")*0.5, labels=FALSE)
axis(2, at=c(seq(0.02,0.09,by=0.01), seq(0.2,0.9,by=0.1), seq(2,9,by=1), seq(20,90,by=10)), tcl=par("tcl")*0.5, labels=FALSE)
#axis(2, at=c(0.1, 1, 10), tcl=par("tcl")*1, labels=FALSE)

plot(overlapykeT~overlapT, 
     ylab="color volume overlap \n(perceptually-corrected, %)", 
     xlab="color volume overlap (%)\n ",
     ylim=plotrange(c(overlapykeT,overlapT), log=FALSE), xlim=plotrange(c(overlapykeT,overlapT), log=FALSE), 
     pch=21, bg=sigpalT, col=NA)
abline(0,1, lty=2)

plot(centdistT~overlapT,
     ylab=' \ncentroid distance (JND)', xlab='color volume overlap (%)\n ',
     ylim=plotrange(centdistT), xlim=plotrange(overlapT, log=FALSE),
     pch=21, bg=sigpalT, col=NA, log='y', yaxt='n', xaxt='n')
axis(1, at=c(0,20,40,60))
axis(2, at=c(0.1, 1, 10), labels=c(0.1, 1, 10))
axis(2, at=c(seq(0.2,0.9, by=0.1), seq(2,9, by=1)), tcl=par("tcl")*0.5, labels=FALSE)
abline(h=1, lty=3)


plot(centdistT~overlapykeT,
     ylab='centroid distance (JND)', xlab='color volume overlap \n(perceptually-corrected, %)',
     ylim=plotrange(centdistT), xlim=plotrange(overlapykeT, log=FALSE),
     pch=21, bg=sigpalT, col=NA, log='y', yaxt='n')
axis(2, at=c(0.1, 1, 10), labels=c(0.1, 1, 10))
axis(2, at=c(seq(0.2,0.9, by=0.1), seq(2,9, by=1)), tcl=par("tcl")*0.5, labels=FALSE)
abline(h=1, lty=3)
dev.off()

```