---
title: "Examples, using real data"
output:
  md_document:
    variant: markdown_github

---

```{r setup, include = FALSE}
require(pavo)
require(scatterplot3d)
require(ggplot2)
require(gridExtra)
require(vegan)
require(lme4)

knitr::opts_chunk$set(echo = TRUE, 
                      fig.path='output/figures/examples/examples_fig')
```

# Possible examples w/ real data 

```{r}
# Stolen from simspt2
adoniscoldist <- function(x){
  dmat <- matrix(0, nrow=length(unique(x$patch1)), ncol=length(unique(x$patch1)))
  rownames(dmat) <- colnames(dmat) <- as.character(unique(x$patch1))
  
  for(i in rownames(dmat))
    for(j in colnames(dmat))
      if(length(x$dS[x$patch1 == i & x$patch2 == j]) != 0)
      dmat[i,j] <- dmat[j,i] <- x$dS[x$patch1 == i & x$patch2 == j]
  
  grouping <- gsub('[0-9]','', rownames(dmat))
  
  adonis(dmat~grouping)
}

# Trichromat space & plot (from pavo)
trispace <- function(vismodeldata){
  
  dat <- vismodeldata
    
  s <- dat[, 1]
  m <- dat[, 2]
  l <- dat[, 3]
  
# cartesian coordinates
    
  x <- (1/sqrt(2)) * (l - m) 
  y <- (sqrt(2)/sqrt(3)) * (s - ((l + m)/2))   
    
# colorimetrics
  r.vec <- sqrt((abs(x))^2 + (abs(y)^2))
  h.theta <- atan2(y, x)
  
  res.p <- data.frame(s, m, l, x, y, h.theta, r.vec, row.names = rownames(dat))
  
  res <- res.p
  
  res
}

triplot <- function(tridata, labels = TRUE, achro = TRUE, achrocol = 'grey', achrosize = 0.8, 
                     cex.labels = 1, out.lwd = 1, out.lcol = 'black', out.lty = 1, ...){ 
  
  arg <- list(...)
  
# Set defaults
  if(is.null(arg$col))
    arg$col <- 'black'
  if(is.null(arg$pch))
    arg$pch <- 19
  if(is.null(arg$type))
    arg$type = 'p'
  if(is.null(arg$xlim))
    arg$xlim <- c(-1/sqrt(2), 1/sqrt(2))
  if(is.null(arg$ylim))
    arg$ylim <- c(-sqrt(2)/(2*(sqrt(3))), sqrt(2)/sqrt(3))
    
# Verticy coordinates  
  vert <- data.frame(x = c(0, -1/sqrt(2), 1/sqrt(2)),
                       y = c(sqrt(2)/sqrt(3), -sqrt(2)/(2*(sqrt(3))), -sqrt(2)/(2*(sqrt(3)))))
  
# Plot
  arg$x <- tridata$x
  arg$y <- tridata$y
  arg$xlab = ' '
  arg$ylab = ' '
  arg$bty = 'n'
  arg$axes = FALSE
  
  do.call(plot, arg)
  
# Add lines 
  segments(vert$x[1], vert$y[1], vert$x[2], vert$y[2], lwd = out.lwd, lty = out.lty, col = out.lcol)
  segments(vert$x[1], vert$y[1], vert$x[3], vert$y[3], lwd = out.lwd, lty = out.lty, col = out.lcol)
  segments(vert$x[2], vert$y[2], vert$x[3], vert$y[3], lwd = out.lwd, lty = out.lty, col = out.lcol)
  
# Origin
  if(isTRUE(achro)){
    points(x = 0, y = 0, pch = 15, col = achrocol, cex = achrosize)
  }
  
# Add text (coloured points better as in tcsplot?)
  if(isTRUE(labels)){
    text('S', x = -0.76, y = -0.39, xpd = TRUE, cex = cex.labels)
    text('M', x = 0, y = 0.88, xpd = TRUE, cex = cex.labels)
    text('L', x = 0.76, y = -0.39, xpd = TRUE, cex = cex.labels)
  }
  
}
```

#### Example 1: Dichromatism. 

Reflectance data from four body regions of male and female _Ctenophorus ornatus_ (Whiting et al. 2015, Biol J Linn Soc). Labium, throat, tongue, and mouth-roof.

**Q:** Which body regions are sexually dichromatic?

Calculate deltaS according to conspecific (tetrachromatic) visual system

```{r message=FALSE, warning=FALSE}

specs <- list(lab = as.rspec(read.csv('data/dichromatism/lab.csv'), interp = FALSE),
              throat = as.rspec(read.csv('data/dichromatism/throat.csv'), interp = FALSE),
              roof = as.rspec(read.csv('data/dichromatism/roof.csv'), interp = FALSE),
              tongue = as.rspec(read.csv('data/dichromatism/tongue.csv'), interp = FALSE))

# Ctenophorus ornatus
liz_vis <- sensmodel(c(360, 440, 493, 571)) 
names(liz_vis) <- c('wl', 'u', 's', 'm', 'l')

models <- lapply(specs, function(x) vismodel(x, visual = liz_vis, relative = FALSE, 
                                             qcatch = "fi", scale = 10000))  # deltaS
models_rel <- lapply(specs, function(x) vismodel(x, visual = liz_vis, relative = TRUE, 
                                                 qcatch = "fi", scale = 10000))  # tcs 

deltaS <- lapply(models, function(x) coldist(x, achro = FALSE, n1 = 1, n2 = 1, 
                                             n3 = 3.5, n4 = 6, v = 0.10))

# To add group labels (because I'm bad at R and I feel bad)
liz_lab <- function(x){
  x$comparison[grepl('F', x$patch1) & grepl('F', x$patch2)] <- 'intra.F'
  x$comparison[grepl('M', x$patch1) & grepl('M', x$patch2)] <- 'intra.M'
  x$comparison[grepl('M', x$patch1) & grepl('F', x$patch2)] <- 'inter'
  x$comparison[grepl('F', x$patch1) & grepl('M', x$patch2)] <- 'inter'
  x
}

# ew
deltaS$lab <- liz_lab(deltaS$lab) 
deltaS$throat <- liz_lab(deltaS$throat)
deltaS$roof <- liz_lab(deltaS$roof)
deltaS$tongue <- liz_lab(deltaS$tongue)


```

Plot 'em

```{r liz_tcs, fig.width=8, fig.height=8}

par(pty="s", mfrow = c(2, 2))

sp3d <- scatterplot3d(suppressWarnings(tcs(models_rel$lab[grepl("M", rownames(models_rel$lab)), ])
                                       [, c('x','y','z')]), pch=19, box=F, main = 'labium')
sp3d$points3d(suppressWarnings(tcs(models_rel$lab[grepl("F", rownames(models_rel$lab)), ])
                               [, c('x','y','z')]), col='red',pch=19)

sp3d <- scatterplot3d(suppressWarnings(tcs(models_rel$throat[grepl("M", rownames(models_rel$throat)), ])
                                       [, c('x','y','z')]), pch=19, box=F, main = 'throat')
sp3d$points3d(suppressWarnings(tcs(models_rel$throat[grepl("F", rownames(models_rel$throat)), ])
                               [, c('x','y','z')]), col='red',pch=19)

sp3d <- scatterplot3d(suppressWarnings(tcs(models_rel$roof[grepl("M", rownames(models_rel$roof)), ])
                                       [, c('x','y','z')]), pch=19, box=F, main = 'roof')
sp3d$points3d(suppressWarnings(tcs(models_rel$roof[grepl("F", rownames(models_rel$roof)), ])
                               [, c('x','y','z')]), col='red',pch=19)

sp3d <- scatterplot3d(suppressWarnings(tcs(models_rel$tongue[grepl("M", rownames(models_rel$tongue)), ])
                                       [, c('x','y','z')]), pch=19, box=F, main = 'tongue')
sp3d$points3d(suppressWarnings(tcs(models_rel$tongue[grepl("F", rownames(models_rel$tongue)), ])
                               [, c('x','y','z')]), col='red',pch=19)
```

```{r liz_deltaplot, fig.width=8, fig.height=8}

p1 <- ggplot(deltaS$lab, aes(x=dS, fill=comparison)) + geom_histogram(bins=50) + 
        facet_grid(comparison~., scales='free_y') + geom_vline(xintercept=1) +
        ggtitle('labial') + theme(legend.position="none")

p2 <- ggplot(deltaS$throat, aes(x=dS, fill=comparison)) + geom_histogram(bins=50) + 
        facet_grid(comparison~., scales='free_y') + geom_vline(xintercept=1) +
        ggtitle('throat') + theme(legend.position="none")

p3 <- ggplot(deltaS$roof, aes(x=dS, fill=comparison)) + geom_histogram(bins=50) + 
        facet_grid(comparison~., scales='free_y') + geom_vline(xintercept=1) +
        ggtitle('roof') + theme(legend.position="none")

p4 <- ggplot(deltaS$tongue, aes(x=dS, fill=comparison)) + geom_histogram(bins=50) + 
        facet_grid(comparison~., scales='free_y') + geom_vline(xintercept=1) +
        ggtitle('tongue') + theme(legend.position="none")

grid.arrange(p1, p2, p3, p4, ncol=2)

```

#### So what's sexually dichromatic?

**Ye olde ways**

Use mean inter-sex distance (eg. Bridge et al. 2008):

```{r}
mean(deltaS$lab[deltaS$lab$comparison == 'inter', ][['dS']])

mean(deltaS$throat[deltaS$throat$comparison == 'inter', ][['dS']])

mean(deltaS$roof[deltaS$roof$comparison == 'inter', ][['dS']])

mean(deltaS$tongue[deltaS$tongue$comparison == 'inter', ][['dS']])

```

Or the maximums?? (eg. Burns & Schultz 2012):

```{r}
max(deltaS$lab[deltaS$lab$comparison == 'inter', ][['dS']])

max(deltaS$throat[deltaS$throat$comparison == 'inter', ][['dS']])

max(deltaS$roof[deltaS$roof$comparison == 'inter', ][['dS']])

max(deltaS$tongue[deltaS$tongue$comparison == 'inter', ][['dS']])
```

Or run some t tests (eg. Igic et al. 2010):

```{r}
t.test(deltaS$lab[deltaS$lab$comparison == 'inter', ][['dS']], mu = 1)

t.test(deltaS$throat[deltaS$throat$comparison == 'inter', ][['dS']], mu = 1)

t.test(deltaS$roof[deltaS$roof$comparison == 'inter', ][['dS']], mu = 1)

t.test(deltaS$tongue[deltaS$tongue$comparison == 'inter', ][['dS']], mu = 1)
```

...etc

**Conclusion**: Dimorphism everywhere!

#### New method

**Step 1:** Run permuational ANOVA (PERMANOVA) on lizard bits to ask if group A is different than group B

Labium

```{r}

adoniscoldist(deltaS$lab)

```

Throat

```{r}

adoniscoldist(deltaS$throat)

```

Mouth-roof

```{r}

adoniscoldist(deltaS$roof)

```

Tongue

```{r}

adoniscoldist(deltaS$tongue)

```

**Conclusion**: labium = distinct, throat = distinct, mouth = nope, tongue = nope.

Much more sensible sensible based on the earlier plots & what the lizards actually look like (in the Whiting et al. paper). Neat.

**Step 2:** Effect sizes for distinct bits (i.e. do they crack the threshold?). Intercept-only mixed-models?

```{r}

deltaS.lab.inter <- deltaS$lab[deltaS$lab$comparison == 'inter', ]

grep('M', deltaS.lab.inter$patch1)
grep('F', deltaS.lab.inter$patch2)

summary(lmer(dS~1+(1|patch1)+(1|patch2), data=deltaS.lab.inter))

```

**95% CI for labium = 4.69, 5.43**

```{r}

deltaS.throat.inter <- deltaS$throat[deltaS$throat$comparison == 'inter', ]

grep('M', deltaS.throat.inter$patch1)
grep('F', deltaS.throat.inter$patch2)

summary(lmer(dS~1+(1|patch1)+(1|patch2), data=deltaS.throat.inter))

```

**95% CI for throat = 5.48, 6.69**

```{r}

rm(deltaS, deltaS.lab.inter, deltaS.throat.inter, models, models_rel, specs, liz_vis, liz_lab)

```

#### Example 2: Mimicry. 

Reflectance data from colour-polymorphic female spiders _Gasteracantha fornicata_, and sympatic flowers from Qld, Australia. (W = white morph, Y = yellow morph, F = flowers)

So three groups, with **two Q's:** 

**(1)** Are spiders actually polymorphic (to prey)?

**(2)** Do spiders (of each morph) resemble sympatric flowers?

```{r}
# specs <- list(spider_w = as.rspec(read.csv('data/mimicry/spiders_white.csv'), interp = FALSE),
#               spider_y = as.rspec(read.csv('data/mimicry/spiders_yellow.csv'), interp = FALSE),
#               flower = as.rspec(read.csv('data/mimicry/flowers.csv'), interp = FALSE))

specs <- as.rspec(read.csv('data/mimicry/flowers_spiders.csv'), interp = FALSE)

```

Calculate deltaS according to bee (trichromat) visual system

```{r}

# Honeybee
bee_vis <- sensmodel(c(350, 440, 540)) 
names(bee_vis) <- c('wl', 's', 'm', 'l')

models <- vismodel(specs, visual = bee_vis, relative = FALSE, 
                                             qcatch = "fi", scale = 10000)  # deltaS
models_rel <- vismodel(specs, visual = bee_vis, relative = TRUE, 
                                                 qcatch = "fi", scale = 10000)  # max triangle
models_tri <- trispace(models_rel)

deltaS <- coldist(models, achro = FALSE)

# Contrast labels
deltaS$comparison[grepl('W_', deltaS$patch1) & grepl('W_', deltaS$patch2)] <- 'intra.W'
deltaS$comparison[grepl('Y_', deltaS$patch1) & grepl('Y_', deltaS$patch2)] <- 'intra.Y'
deltaS$comparison[grepl('F_', deltaS$patch1) & grepl('F_', deltaS$patch2)] <- 'intra.F'
deltaS$comparison[grepl('W_', deltaS$patch1) & grepl('Y_', deltaS$patch2)] <- 'inter.WY'
deltaS$comparison[grepl('Y_', deltaS$patch1) & grepl('W_', deltaS$patch2)] <- 'inter.WY'
deltaS$comparison[grepl('W_', deltaS$patch1) & grepl('F_', deltaS$patch2)] <- 'inter.WF'
deltaS$comparison[grepl('F_', deltaS$patch1) & grepl('W_', deltaS$patch2)] <- 'inter.WF'
deltaS$comparison[grepl('F_', deltaS$patch1) & grepl('Y_', deltaS$patch2)] <- 'inter.YF'
deltaS$comparison[grepl('Y_', deltaS$patch1) & grepl('F_', deltaS$patch2)] <- 'inter.YF'

```

```{r mimic_triplot, fig.width=6, fig.height=6}

triplot(models_tri[grepl("W_", rownames(models_tri)), ], col = 'darkgrey')
points(subset(models_tri[grepl("Y_", rownames(models_tri)), ], select = c('x', 'y')), pch = 19, col = 'yellow')
points(subset(models_tri[grepl("F_", rownames(models_tri)), ], select = c('x', 'y')), pch = 19, col = 'forestgreen')

```

Well that's not terribly helpful. Should actually run it in the hexagon w/ euclidean distances
since it's a bee model. 

```{r mimic_deltaplot, fig.width=6, fig.height=8}

ggplot(deltaS, aes(x=dS, fill=comparison)) + geom_histogram(bins=50) + 
        facet_grid(comparison~., scales='free_y') + geom_vline(xintercept=1) +
        ggtitle('labial') + theme(legend.position="none")

```

Mmm, maybe not the most exciting example. 

...tbc

```{r}

rm(deltaS, models, models_rel, models_tri, specs, bee_vis)

```

#### Example 3: Crypsis. 

Reflectance data from various body regions (H = head, L = left arm, R = right arm, P = prothorax, W = wing, A = abdomen) of 27 female mantids _Pseudomantis albofimbriata_ and 50 background samples (_Lomandra longifolia_, which they pretty much exclusively hang on).

So six groups, one **Q:** Are mantids cryptic? i.e. are all body regions chromaticically 
indistinguishable from their background?

Calculate deltaS according to blue tits 

```{r}

specs <- as.rspec(read.csv('data/crypsis/mantids_bkgs.csv'), lim = c(300, 700))

```

```{r}

models <- vismodel(specs, visual = 'bluetit', relative = FALSE, qcatch = "fi", scale = 10000)  # deltaS
models_rel <- vismodel(specs, visual = 'bluetit', relative = TRUE, qcatch = "fi", scale = 10000, achro = FALSE)  # tcs

deltaS <- coldist(models, achro = FALSE)

```

Plot 'em

```{r mantid_tcs, fig.width=6, fig.height=6}

sp3d <- scatterplot3d(suppressWarnings(tcs(models_rel[grepl("B_", rownames(models_rel)), ])
                                       [, c('x','y','z')]), xlim=c(-0.02,0.02), ylim=c(-0.01,0.028), zlim=c(-0.06,0.01), 
                                        pch=19, box=F, color = 'forestgreen')
sp3d$points3d(suppressWarnings(tcs(models_rel[grepl("H_", rownames(models_rel)), ])
                               [, c('x','y','z')]), col='darkgoldenrod',pch=19)
sp3d$points3d(suppressWarnings(tcs(models_rel[grepl("L_", rownames(models_rel)), ])
                               [, c('x','y','z')]), col='darkgoldenrod1',pch=19)
sp3d$points3d(suppressWarnings(tcs(models_rel[grepl("R_", rownames(models_rel)), ])
                               [, c('x','y','z')]), col='darkgoldenrod2',pch=19)
sp3d$points3d(suppressWarnings(tcs(models_rel[grepl("P_", rownames(models_rel)), ])
                               [, c('x','y','z')]), col='darkgoldenrod3',pch=19)
sp3d$points3d(suppressWarnings(tcs(models_rel[grepl("W_", rownames(models_rel)), ])
                               [, c('x','y','z')]), col='darkgoldenrod4',pch=19)
sp3d$points3d(suppressWarnings(tcs(models_rel[grepl("A_", rownames(models_rel)), ])
                               [, c('x','y','z')]), col='gold1',pch=19)
```

...tbc


```{r}
sessionInfo()
```
