---
title: "Examples, using real data"
output:
  md_document:
    variant: markdown_github

---

```{r setup, include = FALSE}
require(pavo)
require(scatterplot3d)
require(ggplot2)
require(gridExtra)
require(vegan)
require(lme4)

knitr::opts_chunk$set(echo = TRUE, 
                      fig.path='output/figures/examples/examples_fig')
```

# Possible examples w/ real data 

```{r}

source('R/bootstrapcentroiddS.R')
source('R/trispace.R')

# Pared down to just provide the distance matrix (just so I can trace what's going
# on a bit more easily).
distmat <- function(x){
  dmat <- matrix(0, nrow=length(unique(x$patch1)), ncol=length(unique(x$patch1)))
  rownames(dmat) <- colnames(dmat) <- as.character(unique(x$patch1))
  
  for(i in rownames(dmat))
    for(j in colnames(dmat))
      if(length(x$dS[x$patch1 == i & x$patch2 == j]) != 0)
      dmat[i,j] <- dmat[j,i] <- x$dS[x$patch1 == i & x$patch2 == j]

  #grouping <- substring(rownames(dmat), 1, 1)
  
  dmat
  
  #adonis(dmat~grouping)
}

```

## Example 1: Dichromatism. 

Reflectance data from four body regions of male and female _Ctenophorus ornatus_ (Whiting et al. 2015, Biol J Linn Soc). Labium, throat, tongue, and mouth-roof.

**Q:** Which body regions are sexually dichromatic?

Calculate deltaS according to conspecific (tetrachromatic) visual system

```{r message=FALSE, warning=FALSE}

specs <- list(lab = as.rspec(read.csv('data/dichromatism/lab.csv'), interp = FALSE),
              throat = as.rspec(read.csv('data/dichromatism/throat.csv'), interp = FALSE),
              roof = as.rspec(read.csv('data/dichromatism/roof.csv'), interp = FALSE),
              tongue = as.rspec(read.csv('data/dichromatism/tongue.csv'), interp = FALSE))

# Ctenophorus ornatus
liz_vis <- sensmodel(c(360, 440, 493, 571)) 
names(liz_vis) <- c('wl', 'u', 's', 'm', 'l')

models <- lapply(specs, function(x) vismodel(x, visual = liz_vis, relative = FALSE, 
                                             qcatch = "fi", scale = 10000))  # deltaS
models_rel <- lapply(specs, function(x) vismodel(x, visual = liz_vis, relative = TRUE, 
                                                 qcatch = "fi", scale = 10000))  # tcs 

deltaS <- lapply(models, function(x) coldist(x, achro = FALSE, n1 = 1, n2 = 1, 
                                             n3 = 3.5, n4 = 6, v = 0.10, noise = 'neural'))

# Add group labels
liz_lab <- function(x){
  x$comparison[grepl('F', x$patch1) & grepl('F', x$patch2)] <- 'intra.F'
  x$comparison[grepl('M', x$patch1) & grepl('M', x$patch2)] <- 'intra.M'
  x$comparison[grepl('M', x$patch1) & grepl('F', x$patch2)] <- 'inter'
  x$comparison[grepl('F', x$patch1) & grepl('M', x$patch2)] <- 'inter'
  x
}

# ew
deltaS$lab <- liz_lab(deltaS$lab)
deltaS$throat <- liz_lab(deltaS$throat)
deltaS$roof <- liz_lab(deltaS$roof)
deltaS$tongue <- liz_lab(deltaS$tongue)

```

Visualise

```{r liz_tcs, fig.width=8, fig.height=8}

par(pty="s", mfrow = c(2, 2))

sp3d <- scatterplot3d(suppressWarnings(tcs(models_rel$lab[grepl("M", rownames(models_rel$lab)), ])
                                       [, c('x','y','z')]), pch=19, box=F, main = 'labium')
sp3d$points3d(suppressWarnings(tcs(models_rel$lab[grepl("F", rownames(models_rel$lab)), ])
                               [, c('x','y','z')]), col='red',pch=19)

sp3d <- scatterplot3d(suppressWarnings(tcs(models_rel$throat[grepl("M", rownames(models_rel$throat)), ])
                                       [, c('x','y','z')]), pch=19, box=F, main = 'throat')
sp3d$points3d(suppressWarnings(tcs(models_rel$throat[grepl("F", rownames(models_rel$throat)), ])
                               [, c('x','y','z')]), col='red',pch=19)

sp3d <- scatterplot3d(suppressWarnings(tcs(models_rel$roof[grepl("M", rownames(models_rel$roof)), ])
                                       [, c('x','y','z')]), pch=19, box=F, main = 'roof')
sp3d$points3d(suppressWarnings(tcs(models_rel$roof[grepl("F", rownames(models_rel$roof)), ])
                               [, c('x','y','z')]), col='red',pch=19)

sp3d <- scatterplot3d(suppressWarnings(tcs(models_rel$tongue[grepl("M", rownames(models_rel$tongue)), ])
                                       [, c('x','y','z')]), pch=19, box=F, main = 'tongue')
sp3d$points3d(suppressWarnings(tcs(models_rel$tongue[grepl("F", rownames(models_rel$tongue)), ])
                               [, c('x','y','z')]), col='red',pch=19)
```

```{r liz_deltaplot, fig.width=8, fig.height=8}

p1 <- ggplot(deltaS$lab, aes(x=dS, fill=comparison)) + geom_histogram(bins=50) + 
        facet_grid(comparison~., scales='free_y') + geom_vline(xintercept=1) +
        ggtitle('labial') + theme(legend.position="none")

p2 <- ggplot(deltaS$throat, aes(x=dS, fill=comparison)) + geom_histogram(bins=50) + 
        facet_grid(comparison~., scales='free_y') + geom_vline(xintercept=1) +
        ggtitle('throat') + theme(legend.position="none")

p3 <- ggplot(deltaS$roof, aes(x=dS, fill=comparison)) + geom_histogram(bins=50) + 
        facet_grid(comparison~., scales='free_y') + geom_vline(xintercept=1) +
        ggtitle('roof') + theme(legend.position="none")

p4 <- ggplot(deltaS$tongue, aes(x=dS, fill=comparison)) + geom_histogram(bins=50) + 
        facet_grid(comparison~., scales='free_y') + geom_vline(xintercept=1) +
        ggtitle('tongue') + theme(legend.position="none")

grid.arrange(p1, p2, p3, p4, ncol=2)

```

**Step 1:** PERMANOVAs

```{r}

# Setup distance matrices & groupings for each body part
mat <- list(lab = distmat(deltaS$lab),
             throat = distmat(deltaS$throat),
             roof = distmat(deltaS$roof),
             tongue = distmat(deltaS$tongue))
group <- list(lab = substring(rownames(mat$lab), 1, 1),
             throat = substring(rownames(mat$throat), 1, 1),
             roof = substring(rownames(mat$roof), 1, 1),
             tongue = substring(rownames(mat$tongue), 1, 1))

# Labium
adonis(mat$lab ~ group$lab)

# Mouth-roof
adonis(mat$roof ~ group$roof)

# Throat
adonis(mat$throat ~ group$throat)

# Tongue
adonis(mat$tongue ~ group$tongue)

```

**Conclusion**: labium = distinct, throat = distinct, mouth = nope, tongue = nope.

**Step 2:** Effect sizes.

Add grouping variable to raw models, then bootstrap centroids for different patches, as identified in step 1 (labium & throat).

```{r}

# Groups
models$lab$group <- substring(rownames(models$lab), 1, 1)
models$throat$group <- substring(rownames(models$throat), 1, 1)
models$roof$group <- substring(rownames(models$roof), 1, 1)
models$tongue$group <- substring(rownames(models$tongue), 1, 1)

# labium
bootcentroidDS(models$lab[,1:4], models$lab$group, n1 = 1, n2 = 1, n3 = 3.5, n4 = 6, v = 0.10)
# throat
bootcentroidDS(models$throat[,1:4], models$throat$group, n1 = 1, n2 = 1, n3 = 3.5, n4 = 6, v = 0.10)

```

So lab's & throats are statistically distinct, but fall below threshold on average.

```{r}

rm(deltaS, models, models_rel, specs, liz_vis, liz_lab, mat, group)

```

## Example 2: Mimicry. 

Reflectance data from colour-polymorphic female spiders _Gasteracantha fornicata_, and sympatic flowers from Qld, Australia. (W = white morph, Y = yellow morph, F = flowers)

So three groups, with **two Q's:** 

**(1)** Are spiders actually polymorphic (to prey), as naively seems to be the case (for humans)?

**(2)** Do spiders (of each morph) resemble sympatric flowers?

Calculate deltaS (JNDs) according to a honeybee

```{r}

specs <- as.rspec(read.csv('data/mimicry/flowers_spiders.csv'), interp = FALSE)

# Honeybee
bee_vis <- sensmodel(c(350, 440, 540)) 
names(bee_vis) <- c('wl','s', 'm', 'l')

models <- vismodel(specs, visual = bee_vis, relative = FALSE,
                                                 qcatch = "fi", scale = 10000)  # rn
models_rel <- vismodel(specs, visual = bee_vis, relative = TRUE,
                                                 qcatch = "Qi", scale = 10000)  # for plotting
models_tri <- trispace(models_rel)

deltaS <- coldist(models, achro = FALSE)

# Contrast labels
deltaS$comparison[grepl('W_', deltaS$patch1) & grepl('W_', deltaS$patch2)] <- 'intra.W'
deltaS$comparison[grepl('Y_', deltaS$patch1) & grepl('Y_', deltaS$patch2)] <- 'intra.Y'
deltaS$comparison[grepl('F_', deltaS$patch1) & grepl('F_', deltaS$patch2)] <- 'intra.F'
deltaS$comparison[grepl('Y_', deltaS$patch1) & grepl('W_', deltaS$patch2)] <- 'inter.WY'
deltaS$comparison[grepl('W_', deltaS$patch1) & grepl('F_', deltaS$patch2)] <- 'inter.WF'
deltaS$comparison[grepl('Y_', deltaS$patch1) & grepl('F_', deltaS$patch2)] <- 'inter.YF'

```

Visualise.

```{r mimic_triplot, fig.width=6, fig.height=6}

# Max triangle
triplot(models_tri[grepl("F_", rownames(models_tri)), ], col = 'forestgreen')
points(models_tri[grepl("Y_", rownames(models_tri)), ][c('x', 'y')], pch = 19, col = 'darkgoldenrod1')
points(models_tri[grepl("W_", rownames(models_tri)), ][c('x', 'y')], pch = 19, col = 'darkgrey')

```

```{r mimic_deltaplot, fig.width=6, fig.height=6}

ggplot(deltaS, aes(x=dS, fill=comparison)) + geom_histogram(bins=50) + 
        facet_grid(comparison~., scales='free_y') + geom_vline(xintercept=1) +
        theme(legend.position="none")

```

**Step 1:** PERMANOVA
```{r}

# Set up distance matrices & groupings for focal comparisons 
mat <- list(all = distmat(deltaS),
            WY = distmat(subset(deltaS, !(comparison %in% c('intra.F', 'inter.WF', 'inter.YF')))),
            WF = distmat(subset(deltaS, !(comparison %in% c('intra.Y', 'inter.WY', 'inter.YF')))),
            YF = distmat(subset(deltaS, !(comparison %in% c('intra.W', 'inter.WY', 'inter.WF'))))
            )
group <- list(all = substring(rownames(mat$all), 1, 1),
              WY = substring(rownames(mat$WY), 1, 1),
              WF = substring(rownames(mat$WF), 1, 1),
              YF = substring(rownames(mat$YF), 1, 1))

```

Everything combined. Not really required in this situation since we have _a priori_ planned tests.
```{r}

adonis(mat$all ~ group$all)

```

So differences exist.

**Q1:** Are spiders polymorphic?

```{r}

# Spiders only
adonis(mat$WY ~ group$WY)

```

Yep

**Q2:** Are they distinguishable from sympatric flowers?

```{r}

# White morph-vs-flowers
adonis(mat$WF ~ group$WF)

# Yellow morph-vs-flowers
adonis(mat$YF ~ group$YF)

```

White = distinct, yellow = indistinct. Really? That's....unexpected.

**Effect sizes**
```{r}

models$group <- substring(rownames(models), 1, 1)
#bootcentroidDS(models[,1:3], models$group, vis = 'tri')

```

```{r}

rm(deltaS, models, specs, mat, group)

```

## Example 3: Crypsis. 

Reflectance data from various body regions (H = head, L = left arm, R = right arm, P = prothorax, W = wing, A = abdomen) of 27 female mantids _Pseudomantis albofimbriata_ and 50 background samples (_Lomandra longifolia_, which they pretty much exclusively hang on).

So six groups, one **Q:** Are mantids cryptic? i.e. are all body regions chromaticically 
indistinguishable from their background?

Calculate deltaS according to blue tits 

```{r}

specs <- as.rspec(read.csv('data/crypsis/mantids_bkgs.csv'), lim = c(300, 700))

```

```{r}

models <- vismodel(specs, visual = 'bluetit', relative = FALSE, qcatch = "fi", scale = 10000)  # deltaS
models_rel <- vismodel(specs, visual = 'bluetit', relative = TRUE, qcatch = "fi", scale = 10000)  # tcs

deltaS <- coldist(models, achro = FALSE)

# Contrast labels. 

# Intragroup
deltaS$comparison[grepl('H_', deltaS$patch1) & grepl('H_', deltaS$patch2)] <- 'intra.H'
deltaS$comparison[grepl('L_', deltaS$patch1) & grepl('L_', deltaS$patch2)] <- 'intra.L'
deltaS$comparison[grepl('R_', deltaS$patch1) & grepl('R_', deltaS$patch2)] <- 'intra.R'
deltaS$comparison[grepl('P_', deltaS$patch1) & grepl('P_', deltaS$patch2)] <- 'intra.P'
deltaS$comparison[grepl('A_', deltaS$patch1) & grepl('A_', deltaS$patch2)] <- 'intra.A'
deltaS$comparison[grepl('W_', deltaS$patch1) & grepl('W_', deltaS$patch2)] <- 'intra.W'

# Intergroup
deltaS$comparison[grepl('H_', deltaS$patch1) & grepl('B_', deltaS$patch2)] <- 'inter.HB'
deltaS$comparison[grepl('L_', deltaS$patch1) & grepl('B_', deltaS$patch2)] <- 'inter.LB'
deltaS$comparison[grepl('R_', deltaS$patch1) & grepl('B_', deltaS$patch2)] <- 'inter.RB'
deltaS$comparison[grepl('P_', deltaS$patch1) & grepl('B_', deltaS$patch2)] <- 'inter.PB'
deltaS$comparison[grepl('A_', deltaS$patch1) & grepl('B_', deltaS$patch2)] <- 'inter.AB'
deltaS$comparison[grepl('W_', deltaS$patch1) & grepl('B_', deltaS$patch2)] <- 'inter.WB'

deltaS_plot <- deltaS[complete.cases(deltaS),]  # comparisons of interest (for plotting only you fool of a Took)
```

Visualise

```{r crypsis_tcs, fig.width=6, fig.height=6}

sp3d <- scatterplot3d(suppressWarnings(tcs(models_rel[grepl("B_", rownames(models_rel)), ])
                                       [, c('x','y','z')]), xlim=c(-0.02,0.02), ylim=c(-0.01,0.028), zlim=c(-0.06,0.01), 
                                        pch=19, box=F, color = 'forestgreen')
sp3d$points3d(suppressWarnings(tcs(models_rel[grepl("H_", rownames(models_rel)), ])
                               [, c('x','y','z')]), col='darkgoldenrod',pch=19)
sp3d$points3d(suppressWarnings(tcs(models_rel[grepl("L_", rownames(models_rel)), ])
                               [, c('x','y','z')]), col='darkgoldenrod1',pch=19)
sp3d$points3d(suppressWarnings(tcs(models_rel[grepl("R_", rownames(models_rel)), ])
                               [, c('x','y','z')]), col='darkgoldenrod2',pch=19)
sp3d$points3d(suppressWarnings(tcs(models_rel[grepl("P_", rownames(models_rel)), ])
                               [, c('x','y','z')]), col='darkgoldenrod3',pch=19)
sp3d$points3d(suppressWarnings(tcs(models_rel[grepl("W_", rownames(models_rel)), ])
                               [, c('x','y','z')]), col='darkgoldenrod4',pch=19)
sp3d$points3d(suppressWarnings(tcs(models_rel[grepl("A_", rownames(models_rel)), ])
                               [, c('x','y','z')]), col='gold1',pch=19)
```

```{r crypsis_deltaplot, fig.width=6, fig.height=9}

ggplot(deltaS_plot, aes(x=dS, fill=comparison)) + geom_histogram(bins=50) + 
        facet_grid(comparison~., scales='free_y') + geom_vline(xintercept=1) +
        theme(legend.position="none")

```

**Step 1:** PERMANOVA all the things

```{r}

# Set up distance matrices & groupings for focal comparisons 
mat <- list(all = distmat(deltaS)
            #WY = distmat(subset(deltaS, !(comparison %in% c('intra.F', 'inter.WF', 'inter.YF')))),
            )
group <- list(all = substring(rownames(mat$all), 1, 1))
  
adonis(mat$all ~ group$all)

```

**Step 2:** Effect sizes
```{r}

models$group <- substring(rownames(models), 1, 1)
#bootcentroidDS(models[,1:4], models$group)  # Not workin'

```

```{r}
sessionInfo()
```
