---
title: "Example, using data on Sri-Lankan Horned Lizards"
output:
  md_document:
    variant: markdown_github

---

```{r setup, include = FALSE}
require(pavo)
require(scatterplot3d)
require(ggplot2)
require(gridExtra)
require(vegan)

knitr::opts_chunk$set(echo = TRUE, 
                      fig.path='output/figures/lizardeg/lizardeg_fig')
```

# Example w/ real data 

### Reflectance data from several body regions of male and female _Ctenophorus ornatus_ (Whiting et al. 2015, Biol J Linn Soc)

```{r}
# Stolen from simspt2
adoniscoldist <- function(x){
  dmat <- matrix(0, nrow=length(unique(x$patch1)), ncol=length(unique(x$patch1)))
  rownames(dmat) <- colnames(dmat) <- as.character(unique(x$patch1))
  
  for(i in rownames(dmat))
    for(j in colnames(dmat))
      if(length(x$dS[x$patch1 == i & x$patch2 == j]) != 0)
      dmat[i,j] <- dmat[j,i] <- x$dS[x$patch1 == i & x$patch2 == j]
  
  grouping <- gsub('[0-9]','', rownames(dmat))
  
  adonis(dmat~grouping)
  }
```

Calculate deltaS

```{r message=FALSE, warning=FALSE}

specs <- list(lab = as.rspec(read.csv('data/lab.csv'), interp = FALSE),
              throat = as.rspec(read.csv('data/throat.csv'), interp = FALSE),
              roof = as.rspec(read.csv('data/roof.csv'), interp = FALSE),
              tongue = as.rspec(read.csv('data/tongue.csv'), interp = FALSE))

# Ctenophorus ornatus
liz_vis <- sensmodel(c(360, 440, 493, 571)) 
names(liz_vis) <- c('wl', 'u', 's', 'm', 'l')

models <- lapply(specs, function(x) vismodel(x, visual = liz_vis, relative = FALSE, 
                                             qcatch = "fi", scale = 10000))  # deltaS
models_rel <- lapply(specs, function(x) vismodel(x, visual = liz_vis, relative = TRUE, 
                                                 qcatch = "fi", scale = 10000))  # tcs 

deltaS <- lapply(models, function(x) coldist(x, achro = FALSE, n1 = 1, n2 = 1, 
                                             n3 = 3.5, n4 = 6, v = 0.10))

# To add group labels (because I'm bad at R and I feel bad)
comp_lab <- function(x){
  x$comparison[grepl('F', x$patch1) & grepl('F', x$patch2)] <- 'intra.F'
  x$comparison[grepl('M', x$patch1) & grepl('M', x$patch2)] <- 'intra.M'
  x$comparison[grepl('M', x$patch1) & grepl('F', x$patch2)] <- 'inter'
  x$comparison[grepl('F', x$patch1) & grepl('M', x$patch2)] <- 'inter'
  x
}

# ew
deltaS$lab <- comp_lab(deltaS$lab) 
deltaS$throat <- comp_lab(deltaS$throat)
deltaS$roof <- comp_lab(deltaS$roof)
deltaS$tongue <- comp_lab(deltaS$tongue)


```

Plot 'em

```{r tcs, fig.width=8, fig.height=8}

par(pty="s", mfrow = c(2, 2))

sp3d <- scatterplot3d(suppressWarnings(tcs(models_rel$lab[grepl("M", rownames(models_rel$lab)), ])
                                       [, c('x','y','z')]), pch=19, box=F, main = 'labium')
sp3d$points3d(suppressWarnings(tcs(models_rel$lab[grepl("F", rownames(models_rel$lab)), ])
                               [, c('x','y','z')]), col='red',pch=19)

sp3d <- scatterplot3d(suppressWarnings(tcs(models_rel$throat[grepl("M", rownames(models_rel$throat)), ])
                                       [, c('x','y','z')]), pch=19, box=F, main = 'throat')
sp3d$points3d(suppressWarnings(tcs(models_rel$throat[grepl("F", rownames(models_rel$throat)), ])
                               [, c('x','y','z')]), col='red',pch=19)

sp3d <- scatterplot3d(suppressWarnings(tcs(models_rel$roof[grepl("M", rownames(models_rel$roof)), ])
                                       [, c('x','y','z')]), pch=19, box=F, main = 'roof')
sp3d$points3d(suppressWarnings(tcs(models_rel$roof[grepl("F", rownames(models_rel$roof)), ])
                               [, c('x','y','z')]), col='red',pch=19)

sp3d <- scatterplot3d(suppressWarnings(tcs(models_rel$tongue[grepl("M", rownames(models_rel$tongue)), ])
                                       [, c('x','y','z')]), pch=19, box=F, main = 'tongue')
sp3d$points3d(suppressWarnings(tcs(models_rel$tongue[grepl("F", rownames(models_rel$tongue)), ])
                               [, c('x','y','z')]), col='red',pch=19)
```


```{r deltaplot, fig.width=8, fig.height=8}

p1 <- ggplot(deltaS$lab, aes(x=dS, fill=comparison)) + geom_histogram(bins=50) + 
        facet_grid(comparison~., scales='free_y') + geom_vline(xintercept=1) +
        ggtitle('labial') + theme(legend.position="none")

p2 <- ggplot(deltaS$throat, aes(x=dS, fill=comparison)) + geom_histogram(bins=50) + 
        facet_grid(comparison~., scales='free_y') + geom_vline(xintercept=1) +
        ggtitle('throat') + theme(legend.position="none")

p3 <- ggplot(deltaS$roof, aes(x=dS, fill=comparison)) + geom_histogram(bins=50) + 
        facet_grid(comparison~., scales='free_y') + geom_vline(xintercept=1) +
        ggtitle('roof') + theme(legend.position="none")

p4 <- ggplot(deltaS$tongue, aes(x=dS, fill=comparison)) + geom_histogram(bins=50) + 
        facet_grid(comparison~., scales='free_y') + geom_vline(xintercept=1) +
        ggtitle('tongue') + theme(legend.position="none")

grid.arrange(p1, p2, p3, p4, ncol=2)

```

### So what's dimorphic?

**Step 1:** Run permuational ANOVA (PERMANOVA) on lizard bits to ask if group A is different than group B

**Labium**
```{r}

adoniscoldist(deltaS$lab)

```

**Throat**
```{r}

adoniscoldist(deltaS$throat)

```

**Mouth-roof**
```{r}

adoniscoldist(deltaS$roof)

```

**Tongue**
```{r}

adoniscoldist(deltaS$tongue)

```

So this is telling us (?): labium = distinct, throat = distinct, mouth = nope, tongue = nope.

Which seems sensible based on the earlier plots & what the lizards actually look like (in the Whiting et al. paper). Neat.

**Step 2:** Run a linear model to get average within- and between-group distances.

tbc...

```{r}
sessionInfo()
```
