---
title: "Simulations part 2"
output:
  md_document:
    variant: markdown_github

---

```{r setup, include=FALSE}
require(pavo)
require(scatterplot3d)
require(vegan)
require(lme4)

knitr::opts_chunk$set(echo = TRUE, 
                      fig.path='output/cache/simspt2/simspt2_fig', 
                      fig.width=15, fig.height = 10,
                      cache.path = 'output/figures/simspt2/simspt2_cache')

set.seed(6210)
```

## Simulation definitions

- $usml \sim \mathcal{lnN}(ln(\mu), \sigma^2)$
- $\mu_{usml} \sim \mathcal{U}(10^{-6},10^0)$, covariances = 0
- $\sigma^2_{usml} \sim \mathcal{Exp}(\lambda = 10)$ (such that $\overline{\sigma^2} = 0.1$).
- Also, $\sigma^2_{A_{usml}} = \sigma^2_{B_{usml}}$
- difference between group A and group B: $\mu_{A_{usml}} = \mu_{B_{usml}}*X$, where $X \sim \mathcal{N}(\mu=1,\sigma^2=0.01)$ (that is, group B _usml_ should be about 5% different than group A _usml_ 95% of the time)
- $N = 50$

```{r}

simdich <- function(N=50){

  musA <- runif(4, 1e-6,1e0) # vector of means for group A
  #musB <- musA*runif(4, .8, 1.2) # vector of means for group B
  musB <- musA*rnorm(4, 1, 0.01)
  sgsqs <- rexp(4,10) # vector of standard deviations
  
  groupA <- matrix(NA, nrow=N, ncol=4)
  groupA[,1] <- rlnorm(N, meanlog=log(musA[1]), sdlog=sgsqs[1])
  groupA[,2] <- rlnorm(N, meanlog=log(musA[2]), sdlog=sgsqs[2])
  groupA[,3] <- rlnorm(N, meanlog=log(musA[3]), sdlog=sgsqs[3])
  groupA[,4] <- rlnorm(N, meanlog=log(musA[4]), sdlog=sgsqs[4])
  
  groupB <- matrix(NA, nrow=N, ncol=4)
  groupB[,1] <- rlnorm(N, meanlog=log(musB[1]), sdlog=sgsqs[1])
  groupB[,2] <- rlnorm(N, meanlog=log(musB[2]), sdlog=sgsqs[2])
  groupB[,3] <- rlnorm(N, meanlog=log(musB[3]), sdlog=sgsqs[3])
  groupB[,4] <- rlnorm(N, meanlog=log(musB[4]), sdlog=sgsqs[4])
  
  combined <- data.frame(rbind(groupA,groupB))
  
  colnames(combined) <- c('u','s','m', 'l')
  rownames(combined) <- paste(rep(c('gA','gB'), each=N),1:N, sep='')
  
  attr(combined, 'relative') <- FALSE
  
  simpars <- data.frame(rbind(musA, musB, sgsqs))
  colnames(simpars) <- c('u','s','m', 'l')
  rownames(simpars) <- c('muA','muB','ssq')
  attr(combined, 'simpar') <- simpars
  
  combined
  }

```

Simulate 1000 datasets

```{r coldistcache, cache=TRUE}
simulatedata <- lapply(rep(50, 1000), simdich)

simulatecoldist <- parallel::mclapply(simulatedata, function(x) {
  Y <- coldist(x, achro=FALSE)
  Y$comparison <- NA
  Y$comparison[grepl('A', Y$patch1) & grepl('A', Y$patch2)] <- 'intra.A'
  Y$comparison[grepl('B', Y$patch1) & grepl('B', Y$patch2)] <- 'intra.B'
  Y$comparison[grepl('A', Y$patch1) & grepl('B', Y$patch2)] <- 'inter'
  Y
  }, mc.cores=6)
```


Let's see what some of these simulations look like. We can see how similar groups are. This really is a threshold situation. We can also see that simulations do a pretty good job of covering the entire colorspace, as well as a wide range of correlations and variances.

```{r, fig.height=15}
source('R/dichtcp.R')

par(mfrow=c(3,3))

for(i in 1:9) dichtcp(simulatedata[[i]])
```

**Step 1:** Run permuational ANOVA (PERMANOVA) on simulated data to ask if group A is different than group B

```{r adoniscache, cache=TRUE, dependson='coldistcache'}
adonissim <- parallel::mclapply(simulatecoldist, function(x){
  dmat <- matrix(0, nrow=length(unique(x$patch1)), ncol=length(unique(x$patch1)))
  rownames(dmat) <- colnames(dmat) <- as.character(unique(x$patch1))
  
  for(i in rownames(dmat))
    for(j in colnames(dmat))
      if(length(x$dS[x$patch1 == i & x$patch2 == j]) != 0)
      dmat[i,j] <- dmat[j,i] <- x$dS[x$patch1 == i & x$patch2 == j]
  
  grouping <- gsub('[0-9]','', rownames(dmat))
  
  adonis(dmat~grouping)
  }, mc.cores=6)
```


**Step 2:** Run a linear model to get average within- and between-group distances.

```{r lmercache, cache=TRUE, dependson='coldistcache'}

lmesim <- parallel::mclapply(simulatecoldist, function(x) lmer(dS~comparison - 1 + (1|patch1) + (1|patch2), data=x), mc.cores=6)

```

Let's see what our results look like

```{r}
interdist <- unlist(lapply(lmesim, function(x) x@beta[1]))

adonisP <- unlist(lapply(adonissim, function(x) x$aov.tab$'Pr(>F)'[1]))
adonisR2 <- unlist(lapply(adonissim, function(x) x$aov.tab$'R2'[1]))

par(mfrow=c(2,2))

hist(interdist, xlab='mean between-group distance (JND)', breaks=30, col=grey(0.7), main='')
#hist(adonisP)
hist(I(adonisR2*100), xlab='Rsquared from PERMANOVA (%)', breaks=30, col=grey(0.7), main='')
plot(I(adonisR2*100)~interdist, ylab='Rsquared from PERMANOVA (%)', xlab='mean between-group distance (JND)', log='xy', pch=19, col=rgb(0,0,0,0.4))

boxplot(interdist~I(adonisP < 0.05), 
        xlab='PERMANOVA test significant?', ylab='mean between-group distance (JND)', 
        boxwex=0.3, col=grey(0.7), pch=19, log='y')
```

- There is no association between how well groups can be told apart (PERMANOVA R-squared) and the mean between-group distances
- There between-group JND distance is not a good predictor of if the groups can be told apart (i.e. if the PERMANOVA is significant)
- If anything these associations are negative - probably because of the mean-variance relationship in lognormal distributions?


```{r}
sessionInfo()
```
