---
title: "Simulations part 2"
output:
  md_document:
    variant: markdown_github

---

```{r setup, include=FALSE}
require(pavo)
require(scatterplot3d)
require(vegan)
require(lme4)

knitr::opts_chunk$set(echo = TRUE, 
                      fig.path='output/figures/simspt2/simspt2_fig', 
                      fig.height = 10,
                      cache.path = 'output/cache/simspt2/simspt2_cache')

set.seed(6210)
```

## Simulation definitions

- $usml \sim \mathcal{lnN}(ln(\mu), \sigma^2)$
- $\mu_{usml} \sim \mathcal{U}(1, 10)$, covariances = 0
- $\sigma^2_{usml} \sim \mathcal{Exp}(\lambda = 10)$ (such that $\overline{\sigma^2} = 0.1$).
- Also, $\sigma^2_{A_{usml}} = \sigma^2_{B_{usml}}$
- difference between group A and group B: $\mu_{A_{usml}} = \mu_{B_{usml}}*X$, where $X \sim \mathcal{U}(0.95,1.05)$ (that is, group B _usml_ should be up to 5% different than group A _usml_)
- $N_{A} = N_{B} = 50$

```{r}

source('R/dichtcp.R')


simdich <- function(N=50, sgsqsrate=10, multiplier=c(0.95, 1.05)){

  musA <- runif(4, 1e-6, 1e0) # vector of means for group A
  musA <- runif(4, 1, 10)
  #musB <- musA*runif(4, .8, 1.2) # vector of means for group B
  #musB <- musA * rnorm(4, multiplier[1], multiplier[2])
  musB <- musA * runif(4, multiplier[1], multiplier[2])
  sgsqs <- rexp(4, sgsqsrate) # vector of standard deviations
  
  groupA <- matrix(NA, nrow=N, ncol=4)
  groupA[,1] <- rlnorm(N, meanlog=log(musA[1]), sdlog=sgsqs[1])
  groupA[,2] <- rlnorm(N, meanlog=log(musA[2]), sdlog=sgsqs[2])
  groupA[,3] <- rlnorm(N, meanlog=log(musA[3]), sdlog=sgsqs[3])
  groupA[,4] <- rlnorm(N, meanlog=log(musA[4]), sdlog=sgsqs[4])
  
  groupB <- matrix(NA, nrow=N, ncol=4)
  groupB[,1] <- rlnorm(N, meanlog=log(musB[1]), sdlog=sgsqs[1])
  groupB[,2] <- rlnorm(N, meanlog=log(musB[2]), sdlog=sgsqs[2])
  groupB[,3] <- rlnorm(N, meanlog=log(musB[3]), sdlog=sgsqs[3])
  groupB[,4] <- rlnorm(N, meanlog=log(musB[4]), sdlog=sgsqs[4])
  
  combined <- data.frame(rbind(groupA,groupB))
  
  colnames(combined) <- c('u','s','m', 'l')
  rownames(combined) <- paste(rep(c('gA','gB'), each=N),1:N, sep='')
  
  attr(combined, 'relative') <- FALSE
  
  simpars <- data.frame(rbind(musA, musB, sgsqs))
  colnames(simpars) <- c('u','s','m', 'l')
  rownames(simpars) <- c('muA','muB','ssq')
  attr(combined, 'simpar') <- simpars
  
  combined
  }


adoniscoldist <- function(x){
  dmat <- matrix(0, nrow=length(unique(x$patch1)), ncol=length(unique(x$patch1)))
  rownames(dmat) <- colnames(dmat) <- as.character(unique(x$patch1))
  
  for(i in rownames(dmat))
    for(j in colnames(dmat))
      if(length(x$dS[x$patch1 == i & x$patch2 == j]) != 0)
      dmat[i,j] <- dmat[j,i] <- x$dS[x$patch1 == i & x$patch2 == j]
  
  grouping <- gsub('[0-9]','', rownames(dmat))
  
  adonis(dmat~grouping)
  }

```

Simulate 500 datasets

```{r coldistcache, cache=TRUE}
simulatedata <- replicate(500, 
                  simdich(N=50, sgsqsrate=10, multiplier=c(0.95, 1.05)), 
                  simplify=FALSE)

simulatecoldist <- parallel::mclapply(simulatedata, function(x) {
  Y <- coldist(x, achro=FALSE)
  Y$comparison <- NA
  Y$comparison[grepl('A', Y$patch1) & grepl('A', Y$patch2)] <- 'intra.A'
  Y$comparison[grepl('B', Y$patch1) & grepl('B', Y$patch2)] <- 'intra.B'
  Y$comparison[grepl('A', Y$patch1) & grepl('B', Y$patch2)] <- 'inter'
  Y
  }, mc.cores=6)
```


Let's see what some of these simulations look like. We can see how similar groups are. This really is a threshold situation. We can also see that simulations do a pretty good job of covering the entire colorspace, as well as a wide range of correlations and variances.

```{r, echo=FALSE, dependson='coldistcache', fig.width=10}
par(mfrow=c(3,3))

for(i in 1:9) dichtcp(simulatedata[[i]])
```

**Step 1:** Run permuational ANOVA (PERMANOVA) on simulated data to ask if group A is different than group B

```{r adoniscache, cache=TRUE, dependson='coldistcache'}
adonissim <- parallel::mclapply(simulatecoldist, adoniscoldist, mc.cores=6)
```


**Step 2:** Run a linear model to get average within- and between-group distances.

```{r lmercache, cache=TRUE, dependson='coldistcache'}
lmesim <- parallel::mclapply(simulatecoldist, function(x) 
  lmer(dS~comparison - 1 + (1|patch1) + (1|patch2), data=x), 
  mc.cores=6)
```

Let's see what our results look like

```{r, echo=FALSE, dependson=c('coldistcache', 'adoniscache', 'lmercache')}
interdist <- unlist(lapply(lmesim, function(x) x@beta[1]))

adonisP <- unlist(lapply(adonissim, function(x) x$aov.tab$'Pr(>F)'[1]))
adonisR2 <- unlist(lapply(adonissim, function(x) x$aov.tab$'R2'[1]))

intradistA <- unlist(lapply(lmesim, function(x) x@beta[2]))
intradistB <- unlist(lapply(lmesim, function(x) x@beta[3]))

intradistM <- apply(cbind(intradistA,intradistB),1, mean)


par(mfrow=c(3,2))

hist(interdist, xlab='mean distance (JND)', breaks=seq(0,50,by=0.4), col=rgb(1,0,0,0.6), main='', xlim=range(c(interdist,intradistM))* c(0,1.1), ylim=c(0,120))
hist(intradistM, col=rgb(0,0,1,0.6), breaks=seq(0,50,by=0.4), add=TRUE)
legend('topright', fill=c('red','blue'), c('intra','inter'))

plot(interdist~intradistM, ylab='mean between-group distance (JND)', xlab='mean within-group distance (JND)', log='xy', pch=19, col=rgb(0,0,0,0.4))
abline(0,1, lty=3)

plot(I(adonisR2*100)~interdist, ylab='Rsquared from PERMANOVA (%)', xlab='mean between-group distance (JND)', log='xy', pch=19, col=rgb(0,0,0,0.4))

plot(I(adonisR2*100)~intradistM, ylab='Rsquared from PERMANOVA (%)', xlab='mean within-group  distance (JND)', log='xy', pch=19, col=rgb(0,0,0,0.4))

plot(I(adonisR2*100)~I(interdist/intradistM), ylab='Rsquared from PERMANOVA (%)', xlab='between/within group distance ratio', pch=19, col=rgb(0,0,0,0.4))

boxplot(interdist~I(adonisP < 0.05), 
        xlab='PERMANOVA test significant?', ylab='mean between-group distance (JND)', 
        boxwex=0.3, col=grey(0.7), pch=19, log='y')

# boxplot(intradistM~I(adonisP < 0.05), 
#         xlab='PERMANOVA test significant?', ylab='mean within-group distance (JND)', 
#         boxwex=0.3, col=grey(0.7), pch=19, log='y')

rm(adonissim, lmesim, simulatedata, simulatecoldist, interdist, intradistA, intradistB, intradistM, adonisP, adonisR2)
```

- There is no association between how well groups can be told apart (PERMANOVA R-squared) and the mean between-group distances
- There between-group JND distance is not a good predictor of if the groups can be told apart (i.e. if the PERMANOVA is significant)
- If anything these associations are negative - probably because of the mean-variance relationship in lognormal distributions?
- I actually think now that it's because in overall low JNDs, the inter-group variance dominates, but in high JNDs it's essentially just the within-group varince (so lower resolution between-groups).

This is annoying, wasn't really what I was trying to simulate. But makes the point accross... 


Let's try some other simulations.

## Test 1: increase within-group variance

Let's change $\sigma^2_{usml}$ such that $\sigma^2_{usml} \sim \mathcal{Exp}(\lambda = 5)$ (and $\overline{\sigma^2} = 0.2$).


Simulate 500 datasets

```{r coldistcache.t1, cache=TRUE}
simulatedata.t1 <- replicate(500, 
                  simdich(N=50, sgsqsrate=5, multiplier=c(0.95, 1.05)), 
                  simplify=FALSE)

simulatecoldist.t1 <- parallel::mclapply(simulatedata.t1, function(x) {
  Y <- coldist(x, achro=FALSE)
  Y$comparison <- NA
  Y$comparison[grepl('A', Y$patch1) & grepl('A', Y$patch2)] <- 'intra.A'
  Y$comparison[grepl('B', Y$patch1) & grepl('B', Y$patch2)] <- 'intra.B'
  Y$comparison[grepl('A', Y$patch1) & grepl('B', Y$patch2)] <- 'inter'
  Y
  }, mc.cores=6)
```

Groups still overlap a lot but their variation increased a LOT.

```{r, dependson='coldistcache.t1', fig.width=10}
par(mfrow=c(3,3))

for(i in 1:9) dichtcp(simulatedata.t1[[i]])
```

**Step 1:** Run permuational ANOVA (PERMANOVA) on simulated data to ask if group A is different than group B

```{r adoniscache.t1, cache=TRUE, dependson='coldistcache.t1'}
adonissim.t1 <- parallel::mclapply(simulatecoldist.t1, adoniscoldist, mc.cores=6)
```


**Step 2:** Run a linear model to get average within- and between-group distances.

```{r lmercache.t1, cache=TRUE, dependson='coldistcache.t1'}
lmesim.t1 <- parallel::mclapply(simulatecoldist.t1, function(x) 
  lmer(dS~comparison - 1 + (1|patch1) + (1|patch2), data=x), 
  mc.cores=6)
```

What do the results look like in this case?

```{r, echo=FALSE, dependson=c('coldistcache.t1', 'adoniscache.t1', 'lmercache.t1')}
interdist.t1 <- unlist(lapply(lmesim.t1, function(x) x@beta[1]))

adonisP.t1 <- unlist(lapply(adonissim.t1, function(x) x$aov.tab$'Pr(>F)'[1]))
adonisR2.t1 <- unlist(lapply(adonissim.t1, function(x) x$aov.tab$'R2'[1]))

intradistA.t1 <- unlist(lapply(lmesim.t1, function(x) x@beta[2]))
intradistB.t1 <- unlist(lapply(lmesim.t1, function(x) x@beta[3]))

intradistM.t1 <- apply(cbind(intradistA.t1,intradistB.t1),1, mean)


par(mfrow=c(3,2))

hist(interdist.t1, xlab='mean distance (JND)', breaks=seq(0,100,by=1), col=rgb(1,0,0,0.6), main='', xlim=range(c(interdist.t1,intradistM.t1))* c(0,1.1), ylim=c(0,120))
hist(intradistM.t1, col=rgb(0,0,1,0.6), breaks=seq(0,100,by=1), add=TRUE)
legend('topright', fill=c('red','blue'), c('intra','inter'))

plot(interdist.t1~intradistM.t1, ylab='mean between-group distance (JND)', xlab='mean within-group distance (JND)', log='xy', pch=19, col=rgb(0,0,0,0.4))
abline(0,1, lty=3)

plot(I(adonisR2.t1*100)~interdist.t1, ylab='Rsquared from PERMANOVA (%)', xlab='mean between-group distance (JND)', log='xy', pch=19, col=rgb(0,0,0,0.4))

plot(I(adonisR2.t1*100)~intradistM.t1, ylab='Rsquared from PERMANOVA (%)', xlab='mean within-group  distance (JND)', log='xy', pch=19, col=rgb(0,0,0,0.4))

plot(I(adonisR2.t1*100)~I(interdist.t1/intradistM.t1), ylab='Rsquared from PERMANOVA (%)', xlab='between/within group distance ratio', pch=19, col=rgb(0,0,0,0.4))

boxplot(interdist.t1~I(adonisP.t1 < 0.05), 
        xlab='PERMANOVA test significant?', ylab='mean between-group distance (JND)', 
        boxwex=0.3, col=grey(0.7), pch=19, log='y')

# boxplot(intradistM.t1~I(adonisP.t1 < 0.05), 
#         xlab='PERMANOVA test significant?', ylab='mean within-group distance (JND)', 
#         boxwex=0.3, col=grey(0.7), pch=19, log='y')

rm(adonissim.t1, lmesim.t1, simulatedata.t1, simulatecoldist.t1, interdist.t1, intradistA.t1, intradistB.t1, intradistM.t1, adonisP.t1, adonisR2.t1)
```

We see the same results. Note that the between-group distance also increased tenfold, but that's just because of the within-group increase (they're essentially sampled from the same population!).

## Test 2: increase between-group distance

Now let's change the multiplier for the difference between group A and group B such that $X \sim \mathcal{U}(0.6,1.4)$ (that is, group B _usml_ should be up to 40% different than group A _usml_)


Simulate 500 datasets

```{r coldistcache.t2, cache=TRUE}
simulatedata.t2 <- replicate(500, 
                  simdich(N=50, sgsqsrate=10, multiplier=c(0.6, 1.4)), 
                  simplify=FALSE)

simulatecoldist.t2 <- parallel::mclapply(simulatedata.t2, function(x) {
  Y <- coldist(x, achro=FALSE)
  Y$comparison <- NA
  Y$comparison[grepl('A', Y$patch1) & grepl('A', Y$patch2)] <- 'intra.A'
  Y$comparison[grepl('B', Y$patch1) & grepl('B', Y$patch2)] <- 'intra.B'
  Y$comparison[grepl('A', Y$patch1) & grepl('B', Y$patch2)] <- 'inter'
  Y
  }, mc.cores=6)
```

Groups still overlap a lot but their variation increased a LOT.

```{r, dependson='coldistcache.t2', fig.width=10}
par(mfrow=c(3,3))

for(i in 1:9) dichtcp(simulatedata.t2[[i]])
```

**Step 1:** Run permuational ANOVA (PERMANOVA) on simulated data to ask if group A is different than group B

```{r adoniscache.t2, cache=TRUE, dependson='coldistcache.t2'}
adonissim.t2 <- parallel::mclapply(simulatecoldist.t2, adoniscoldist, mc.cores=6)
```


**Step 2:** Run a linear model to get average within- and between-group distances.

```{r lmercache.t2, cache=TRUE, dependson='coldistcache.t2'}
lmesim.t2 <- parallel::mclapply(simulatecoldist.t2, function(x) 
  lmer(dS~comparison - 1 + (1|patch1) + (1|patch2), data=x), 
  mc.cores=6)
```

What do the results look like in this case?

```{r, echo=FALSE, dependson=c('coldistcache.t2', 'adoniscache.t2', 'lmercache.t2')}
interdist.t2 <- unlist(lapply(lmesim.t2, function(x) x@beta[1]))

adonisP.t2 <- unlist(lapply(adonissim.t2, function(x) x$aov.tab$'Pr(>F)'[1]))
adonisR2.t2 <- unlist(lapply(adonissim.t2, function(x) x$aov.tab$'R2'[1]))

intradistA.t2 <- unlist(lapply(lmesim.t2, function(x) x@beta[2]))
intradistB.t2 <- unlist(lapply(lmesim.t2, function(x) x@beta[3]))

intradistM.t2 <- apply(cbind(intradistA.t2,intradistB.t2),1, mean)


par(mfrow=c(3,2))

hist(interdist.t2, xlab='mean distance (JND)', breaks=seq(0,50,by=0.4), col=rgb(1,0,0,0.6), main='', xlim=range(c(interdist.t2,intradistM.t2))* c(0,1.1), ylim=c(0,120))
hist(intradistM.t2, col=rgb(0,0,1,0.6), breaks=seq(0,50,by=0.4), add=TRUE)
legend('topright', fill=c('red','blue'), c('intra','inter'))

plot(interdist.t2~intradistM.t2, ylab='mean between-group distance (JND)', xlab='mean within-group distance (JND)', log='xy', pch=19, col=rgb(0,0,0,0.4))
abline(0,1, lty=3)

plot(I(adonisR2.t2*100)~interdist.t2, ylab='Rsquared from PERMANOVA (%)', xlab='mean between-group distance (JND)', log='xy', pch=19, col=rgb(0,0,0,0.4))

plot(I(adonisR2.t2*100)~intradistM.t2, ylab='Rsquared from PERMANOVA (%)', xlab='mean within-group  distance (JND)', log='xy', pch=19, col=rgb(0,0,0,0.4))

plot(I(adonisR2.t2*100)~I(interdist.t2/intradistM.t2), ylab='Rsquared from PERMANOVA (%)', xlab='between/within group distance ratio', pch=19, col=rgb(0,0,0,0.4))

boxplot(interdist.t2~I(adonisP.t2 < 0.05), 
        xlab='PERMANOVA test significant?', ylab='mean between-group distance (JND)', 
        boxwex=0.3, col=grey(0.7), pch=19, log='y')

# boxplot(intradistM.t2~I(adonisP.t2 < 0.05), 
#        xlab='PERMANOVA test significant?', ylab='mean within-group distance (JND)', 
#        boxwex=0.3, col=grey(0.7), pch=19, log='y')

rm(adonissim.t2, lmesim.t2, simulatedata.t2, simulatecoldist.t2, interdist.t2, intradistA.t2, intradistB.t2, intradistM.t2, adonisP.t2, adonisR2.t2)
```

What is this sorcery? In this case significant results even have *lower* deltaS! We can see that's because high values of between-group distance also have high values of within-group distance in our simulations.

```{r}
sessionInfo()
```
