---
title: "Simulation example"
output:
  html_document: default
---

# Simulation

Here we'll attempt to show, using simulations, that using mean deltaS to estimate the difference between two groups might not be appropriate.

```{r}
require(pavo)
require(scatterplot3d)
require(ggplot2)

sessionInfo()
```


### Example 1: low intra-group variability, low inter-group distance

```{r}
# step 1: generate data

set.seed(1982)

# we'll just consider usml are uncorrelated for this example

# we'll generate data from a lognormal distribution to avoid negative values
# and a variance (on the log scale) of 0.002

groupA <- data.frame(
  u = rlnorm(30, meanlog=log(0.1), sdlog=sqrt(0.002)),
  s = rlnorm(30, meanlog=log(0.1), sdlog=sqrt(0.002)),
  m = rlnorm(30, meanlog=log(0.3), sdlog=sqrt(0.002)),
  l = rlnorm(30, meanlog=log(0.7), sdlog=sqrt(0.002))
)

groupB <- data.frame(
  u = rlnorm(30, meanlog=log(0.1), sdlog=sqrt(0.002)),
  s = rlnorm(30, meanlog=log(0.1), sdlog=sqrt(0.002)),
  m = rlnorm(30, meanlog=log(0.3), sdlog=sqrt(0.002)),
  l = rlnorm(30, meanlog=log(0.7), sdlog=sqrt(0.002))
)


colnames(groupA) <- colnames(groupB) <- c('u','s','m', 'l')
attr(groupA, 'relative') <- attr(groupB, 'relative') <- FALSE

par(pty="s")
sp3d <- scatterplot3d(suppressWarnings(tcs(groupA)[, c('x','y','z')]), pch=19,
                      xlim=c(0.26,0.38), ylim=c(-0.17,0.024), zlim=c(-0.2,-0.14), box=F)
sp3d$points3d(suppressWarnings(tcs(groupB)[, c('x','y','z')]), col='red',pch=19)

```

Note that although USML were simulated uncorrelated, XYZ are correlated. 

Calculate deltaS

```{r}
rownames(groupA) <- paste('gA',1:30,sep='')
rownames(groupB) <- paste('gB',1:30,sep='')

alldat <- rbind(groupA, groupB)

deltaS <- coldist(alldat, achro=FALSE)

deltaS$comparison <- NA

deltaS$comparison[grepl('A', deltaS$patch1) & grepl('A', deltaS$patch2)] <- 'intra.A'
deltaS$comparison[grepl('B', deltaS$patch1) & grepl('B', deltaS$patch2)] <- 'intra.B'
deltaS$comparison[grepl('A', deltaS$patch1) & grepl('B', deltaS$patch2)] <- 'inter'

ggplot(deltaS, aes(x=dS, fill=comparison)) + geom_histogram(bins=50) + 
  facet_grid(comparison~., scales='free_y') + geom_vline(xintercept=1) +
  theme(legend.position="none")

```



### Example 2: High intra-group variability, low inter-group distance

```{r}
# step 1: generate data

set.seed(1982)

# we'll just consider usml are uncorrelated for this example

# we'll generate data from a lognormal distribution to avoid negative values
# and a variance (on the log scale) of 0.002

groupA <- data.frame(
  u = rlnorm(30, meanlog=log(0.1), sdlog=sqrt(0.02)),
  s = rlnorm(30, meanlog=log(0.1), sdlog=sqrt(0.02)),
  m = rlnorm(30, meanlog=log(0.3), sdlog=sqrt(0.02)),
  l = rlnorm(30, meanlog=log(0.7), sdlog=sqrt(0.02))
)

groupB <- data.frame(
  u = rlnorm(30, meanlog=log(0.1), sdlog=sqrt(0.02)),
  s = rlnorm(30, meanlog=log(0.1), sdlog=sqrt(0.02)),
  m = rlnorm(30, meanlog=log(0.3), sdlog=sqrt(0.02)),
  l = rlnorm(30, meanlog=log(0.7), sdlog=sqrt(0.02))
)


colnames(groupA) <- colnames(groupB) <- c('u','s','m', 'l')
attr(groupA, 'relative') <- attr(groupB, 'relative') <- FALSE

par(pty="s")
sp3d <- scatterplot3d(suppressWarnings(tcs(groupA)[, c('x','y','z')]), pch=19,
                      xlim=c(0.26,0.38), ylim=c(-0.17,0.024), zlim=c(-0.2,-0.14), box=F)
sp3d$points3d(suppressWarnings(tcs(groupB)[, c('x','y','z')]), col='red',pch=19)

```

Note that although USML were simulated uncorrelated, XYZ are correlated. 

Calculate deltaS

```{r}
rownames(groupA) <- paste('gA',1:30,sep='')
rownames(groupB) <- paste('gB',1:30,sep='')

alldat <- rbind(groupA, groupB)

deltaS <- coldist(alldat, achro=FALSE)

deltaS$comparison <- NA

deltaS$comparison[grepl('A', deltaS$patch1) & grepl('A', deltaS$patch2)] <- 'intra.A'
deltaS$comparison[grepl('B', deltaS$patch1) & grepl('B', deltaS$patch2)] <- 'intra.B'
deltaS$comparison[grepl('A', deltaS$patch1) & grepl('B', deltaS$patch2)] <- 'inter'

ggplot(deltaS, aes(x=dS, fill=comparison)) + geom_histogram(bins=50) + 
  facet_grid(comparison~., scales='free_y') + geom_vline(xintercept=1) +
  theme(legend.position="none")

```


#### Note that in both cases there is no dichromatism, but regular analyses would consider the second case dichromatic.

## How to test this?
I think something from the likes of the adonis function in vegan, but I don't know how to use a custom distance matrix with it yet.